---
title: "HFVI"
author: "Annika Kunz"
date: "2024-01-11"
output: html_document
---

This code runs the calculation of the HFVI applied to the Mahama Case Study. Data from the Mahama Camp are therefore used as inputs to create the individual indicators raster layer for subsequent spatial overlay, using the HFVI equation with the FAHP weights (obtained from running AHP.Rmd and FAHP.Rmd). Further spatial analysis (spatial autocorrelation and indicator correlation matrix) is performed followed by sensitivity and uncertainty analysis with the implementation of the FAHP-OAT Method (using the fuzzy ranges from the FAHP.Rmd file). Finally, visualization of the HFVI and uncertainty maps is performed.

This code can be adjusted for other case studys, changing the input data for the individual indicator layers. 

# Load Libraries

```{r setup, include=FALSE, warning = F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# Install/load packages
## Default repository
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org" 
       options(repos = r)
})

check_pkg <- function(x)
  {
    if (!require(x, character.only = TRUE))
    {
      install.packages(x, dep = TRUE)
        if(!require(x, character.only = TRUE)) stop("Package not found")
    }
}

check_pkg("sf")
check_pkg("httr")
check_pkg("rgdal")
check_pkg("ggplot2")
check_pkg("plotly")
check_pkg("tidyverse")
check_pkg("leaflet")
check_pkg("dplyr")
check_pkg("XML")
check_pkg("mapview")
check_pkg("lubridate")
check_pkg("dplyr")
check_pkg("readr") 
check_pkg("tmap")
check_pkg("spatstat") 
check_pkg("devtools")
check_pkg("stringr")
check_pkg("zoo")
check_pkg("RColorBrewer")
check_pkg("spatstat")
check_pkg("here")
check_pkg("maptools")
check_pkg("arrow") #feather.R
check_pkg("raster")
check_pkg("terra")
check_pkg("tmaptools")
check_pkg("stars")
check_pkg("gridExtra")
check_pkg("spatialEco")
check_pkg("corrplot")
check_pkg("ggpattern")

# check_pkg("GD")
# check_pkg("IDSA") # for fuzzy overlay
#check_pkg("fsr")

# install.packages("devtools")
# library(devtools)
# install_github("accarniel/fsr")
library("fsr")

# Install SpatMCDA from GitHub
library(devtools)
devtools::install_github('Neaop/SpatMCDA', build_vignettes = TRUE)
library(SpatMCDA)
```

```{r Projections}
# Projection strings for the Swiss LV03 & LV95 CRS and WGS84 CRS, respectively
crs_wgs84 <- 4326
crs_utm <- 32736 # WGS 84 / UTM zone 36S EPSG:32736
```

```{r set working directory}
#setwd("~/Documents/Uni/Master/Masterarbeit/R")
```

# Initialization of project paths
```{r init_paths, warning = FALSE, message = FALSE, fig.align = "center"}
library(here)
here::i_am("HFVI.Rmd")
knitr::opts_knit$set(root.dir=here("R/"))

dataFolder   <- here::here("data")   # Data folder
RFolder      <- here::here()         # RScript folder (i.e. where project file resides)
figureFolder <- here::here("figs")   # Figure folder
```

```{r Color Brewer}
#palette_explorer()
```


# Read Data

- OSM data: https://download.geofabrik.de/africa/rwanda.html 
- Mahama refugee camp boundary: hand drawn in the UNHCR GIS tool based on satellite image > ADJUST FOR OTHER CAMPS, PREPOCESSING IS NECESSARY (e.g. using ArcGIS Pro)
- Google buildings: https://sites.research.google/open-buildings/#dataformat 
- Global land over data 30-m spatial resultion: https://glad.umd.edu/dataset/GLCLUC2020 

```{r Read data}
# Read RWA OSM shapefiles and transfrom to crs to EPSG:32736
data_osm_build <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_buildings_a_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_landuse <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_landuse_a_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_natural <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_natural_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_places <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_places_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_pofw <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_pofw_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_pois <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_pois_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_railways <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_railways_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_roads <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_roads_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_traffic <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_traffic_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_transport <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_transport_free_1.shp") %>%
  st_transform(crs = crs_utm)
data_osm_waterways <- read_sf("./data/osm/rwanda_osm.shp/gis_osm_waterways_free_1.shp") %>%
  st_transform(crs = crs_utm)

# Google buildings
# data_buildings <- read.csv(".data/buildings/19d_buildings.csv")
```

```{r Feature classes created in ArcGIS Pro}
# Load shapefiles of geometries created in ArcGIS Pro based on UNHCR camp situation map 2020

# Read Mahama camp boundary 
# mahama_camp_boundary <- read_sf("./data/geometries/boundary_Mahama/Mahama_camp_boundary.shp")
mahama_boundary <- read_sf("./data/geometries/boundary_extended/Mahama_boundary_extended.shp") %>% # extended camp boundary 
  st_transform(crs = crs_utm)

summary(mahama_boundary)

# Read facilities 
mahama_facilities <- read_sf("./data/geometries/facilities_UNHCR/Mahama_Facilities.shp") %>% subset(., select=-c(Name)) %>% # Drop empty column 
  st_transform(crs = crs_utm)

# Read land use 
landuse <- read_sf("./data/geometries/landuse_UNHCR/landuse.shp") %>%
  st_transform(crs = crs_utm)

# Shelter
# Filtering operations and correction in ArcGIS Pro 
mahama_shelter <- st_read("./data/buildings/mahama_shelter/shelter_final.shp") %>% subset(., select=-c(6:9)) %>%
  st_transform(crs = crs_utm)

# Sanitary units
mahama_sanitary <- st_read("./data/buildings/sanitary_units/basic_services.shp")  %>% subset(., select=-c(6:9)) %>%
  st_transform(crs = crs_utm)

# Vulnerable Groups
mahama_vulnerable_groups <- st_read("./data/vulnerable_groups/vul_groups_def.shp") %>%
  st_transform(crs = crs_utm)

# Critical Infrastructure
mahama_critical_infra <- st_read("./data/critical_infrastructure/ci_def.shp") %>%
  st_transform(crs = crs_utm)

# Shelter Type 
mahama_shelter_type <- st_read("./data/buildings/shelter_type/shelter_type_def.shp") %>%
  st_transform(crs = crs_utm)

# Drainage System
mahama_drainage <- st_read("./data/geometries/drainage_mahama/drainage.shp") %>%
  st_transform(crs = crs_utm)

# Land use
mahama_land_cover <- st_read("./data/land_use/land_cover.shp") %>%
  st_transform(crs = crs_utm)
```

Global land over data 30-m spatial resultion: https://glad.umd.edu/dataset/GLCLUC2020 

```{r Land Cover GeoTIFF}
# # Load landuse geotif 
# 
# GLCLUC2020 <- raster("./data/land_use/00N_030E.tif")
# 
# # Define target projection (WGS 84 / UTM zone 36S)
# target_crs <- crs("+proj=utm +zone=36 +south +datum=WGS84 +units=m +no_defs")
# 
# GLCLUC2020_projected <- projectRaster(GLCLUC2020, crs = target_crs)
```


```{r Create intersection shapefile of Mahama buildings}
# Google Buildings data
# class(data_buildings)
# data_buildings_sf <- st_as_sf(data_buildings)
# head(data_buildings_sf)

# mah_buildings <- st_intersection(data_buildings_sf, mahama_camp_boundary)

# Write shapefile to redgruce the data storage size of the building dataset to reduce run time for later use
# st_write(mah_buildings, "data/buildings/mahama_buildings.shp")

# Read created Mahama buidlings intersection shapefile 
mah_buildings <- st_read("./data/buildings/mahama_buildings/mahama_buildings.shp") %>%
  st_transform(crs = crs_utm)
```
# Data Preprocessing

```{r Camp boundary intersection}
# Clip to camp extent

# OSM data 
mah_osm_build <- st_intersection(data_osm_build, mahama_boundary)
mah_osm_landuse <- st_intersection(data_osm_landuse, mahama_boundary)
mah_osm_natural <- st_intersection(data_osm_natural, mahama_boundary)
mah_osm_places <- st_intersection(data_osm_places, mahama_boundary)
mah_osm_pofw <- st_intersection(data_osm_pofw, mahama_boundary)
mah_osm_pois <- st_intersection(data_osm_pois, mahama_boundary)
mah_osm_railways <- st_intersection(data_osm_railways, mahama_boundary)
mah_osm_roads <- st_intersection(data_osm_roads, mahama_boundary)
mah_osm_traffic <- st_intersection(data_osm_traffic, mahama_boundary)
mah_osm_transport <- st_intersection(data_osm_transport, mahama_boundary)
mah_osm_waterways <- st_intersection(data_osm_waterways, mahama_boundary)

# ArcGIS corrected data
mah_buildings <- st_intersection(mah_buildings, mahama_boundary) 
mah_facilities <- st_intersection(mahama_facilities, mahama_boundary)
mah_shelter <- st_intersection(mahama_shelter, mahama_boundary)
mah_sanitary <- st_intersection(mahama_sanitary, mahama_boundary)
mah_landuse <-st_intersection(landuse, mahama_boundary)
mah_vul_groups <- mahama_vulnerable_groups
mah_ci <- st_intersection(mahama_critical_infra, mahama_boundary)
mah_shelter_type <- st_intersection(mahama_shelter_type, mahama_boundary)
mah_drainage <- st_intersection(mahama_drainage, mahama_boundary)
mah_land_cover <- st_intersection(mahama_land_cover, mahama_boundary)
```

```{r Plot the data}
tmap_mode("plot")

# Create a tmap object with multiple layers
tm_osm_mahama <- tm_shape(mahama_boundary) +
  tm_borders(col = "cyan", lwd = 2, alpha = 0.8) +  # Boundary
  
  # OSM roads
  tm_shape(mah_osm_roads) +
  tm_lines(lwd = 1, 
           col = "fclass",
           palette = "Set1") +  # OSM Roads
  
  # Drainage System
  tm_shape(mah_drainage) +
  tm_lines(lwd = 2, col = "#1f78b4") +
  tm_add_legend(type = "line", labels = "Drainage System", lwd = 2, col = "#1f78b4", title = "Critical Infrastructure") +

  # Sanitary
  tm_shape(mah_sanitary) +
  tm_fill(col = "#fbb4ae", alpha = 0.8, legend.show = TRUE) +
  tm_borders(col = "black", lwd = 0.5) +  # Add black border
  tm_add_legend(type = "fill", labels = "Sanitary Units", col = "#fbb4ae") +  # Sanitary
  
  # Critical Building Infrastructure
  tm_shape(mah_ci) +
  tm_fill("#ccebc5", alpha = 0.5) +
  tm_borders(col = "black", lwd = 0.5) +
  tm_add_legend(type = "fill", labels = "Critical Building Infrastructure", col = "#ccebc5") +
  
  # UNHCR facilities
  tm_shape(mah_facilities) + 
  tm_fill(col = "Facility", alpha = 0.8, palette = "Pastel2") +  # Facility
  
  # Land use
  tm_shape(mah_landuse) +
  tm_fill("#ffffcc", alpha = 0.5) +
  tm_borders(col = "black", lwd = 0.5) +
  tm_add_legend(type = "fill", labels = "Agriculture", col = "#ffffcc", title = "Land Use") +
  
  # Vulnerable Groups
  tm_shape(mah_vul_groups) +
  tm_fill("#fed9a6", alpha = 0.5) +
  tm_borders(col = "black", lwd = 0.5) +
  tm_add_legend(type = "fill", labels = "Vulnerable Groups", col = "#fed9a6", title = "Presence of vulnerable groups") +
  
  # Google Buildings 
  tm_shape(mah_buildings) +
  tm_fill("#AEC7E8", alpha = 0.8) +
  tm_borders(col = "black", lwd = 0.1) +
  tm_add_legend(type = "fill", labels = "Buildings", col = "#AEC7E8", title = "Buildings") +
  
  # Shelter Type
  tm_shape(mah_shelter_type) +
  tm_fill("#decbe4", alpha = 0.5) +
  tm_borders(col = "black", lwd = 0.5) +
  tm_add_legend(type = "fill", labels = "Shelter Type", col = "#decbe4") +

  
  # Basemap options
  tm_basemap("Esri.WorldTopoMap") +  # Basemap 1
  tm_basemap("Esri.WorldImagery") +  # Basemap 2
  tm_basemap("Stamen.TonerLite") +  # Basemap 3
  tm_basemap("OpenStreetMap.Mapnik") +  # Basemap 4
  
  tm_layout(legend.outside = TRUE, frame = FALSE) +  # Place legend outside the plot

  tm_scale_bar(position = c("left", "bottom")) +  # Add scale bar
  tm_compass(type="arrow", position=c(0, .2))  # Add compass

# Display the map
tm_view(tm_osm_mahama)

# Save tmap 
tmap_save(tm_osm_mahama, "vector_map.png", height = 7)
```

In order to get a new dataframe which only contains the facilities, the mapped facility areas are intersected with the building dataset 
```{r Filter shelters}
# First, create multipolygon 
mah_buildings_sfc <- st_combine(mah_buildings)
mah_facilities_sfc <- st_combine(mah_facilities)

# Calculate the intersection
facility_buildings <- st_intersection(mah_buildings, mah_facilities) #all buildings which are located in a mapped facility area

mapview(facility_buildings)

mapview(mah_shelter)

# the remaining intersections (st_differences) were calculated in ArcGIS Pro due to easier computation and instant visual verification and correction
# Translate into code for reproducibility
```

## Attention: Don't run twice!!
```{r Correct shelters}
# Define the IDs to drop
ids_to_drop <- c(985, 4958, 983, 3037, 19, 1972, 5989, 6722, 1468)

# Drop rows with the specified indices
mah_shelter <- mah_shelter %>%
  slice(-ids_to_drop)
```

```{r Plot shelter and facility buildings}
# Check
tmap_mode("plot")

basemap_imagery <-  tm_basemap("Esri.WorldTopoMap") 

# Create a tmap object with two layers
tm_shelter_mahama <- tm_shape(mahama_boundary) +
  tm_borders(col = "cyan", lwd = 2, alpha = 0.8) +
  
  #UNHCR facilities
  tm_shape(facility_buildings) +
  tm_fill(col = "Facility", alpha = 0.8) +

  # Shelter
  tm_shape(mah_shelter) +
  tm_fill("#808080", alpha = 0.8) +
  
  # Sanitary units
  tm_shape(mah_sanitary) +
  tm_fill("blue", alpha = 0.8) +
  
  basemap_imagery

# Display the map
tm_view(tm_shelter_mahama)

# Export shapefile
#st_write(facility_buildings, "mah_facility.shp")
```

Determine grid size;
1. Determine the latitude of the Mahama Refugee Camp.
2. Calculate the approximate size of one degree of latitude at that latitude using the formula.
3. Convert the size from degrees to kilometers.

```{r Calculate grid cell size for specific location}
# # Latitude of the Mahama Refugee Camp (approximate)
# # 30.8454002°E 2.3038914°S 
# 
# # Latitude of the Mahama Refugee Camp (approximate)
# 
# latitude_mahama <- -2.3038914 
# 
# # Calculate the length of one degree of latitude at the given latitude (in meters)
# length_one_degree_latitude_m <- 111132.92 - 559.82 * cos(2 * latitude_mahama * pi / 180) + 1.175 * cos(4 * latitude_mahama * pi / 180)
# 
# # Calculate the size of one meter in degrees latitude
# one_meter_in_degrees_latitude <- 1 / length_one_degree_latitude_m
# 
# # Calculate the size of the grid cells in degrees -> insert desired grid cell size -> 30m grid 
# grid_cell_size_degrees <- 30 * one_meter_in_degrees_latitude
# 
# # Print the size of the grid cells in degrees
# print(grid_cell_size_degrees)
```

```{r Bounding box}
# Define grid cell size
grid_cell_size <- 30

# Calculate the bounding box of mahama_boundary
bbox_mahama <- st_bbox(mahama_boundary)

# Define the upper left corner coordinates with additional grid cell sizes in both x and y directions
upper_left_corner <- c(x = bbox_mahama[1] - (2 * grid_cell_size), y = bbox_mahama[4] + (4 * grid_cell_size))

# Define the dimensions of the rectangle in meters with additional grid cell sizes
x_length <- (55 + 4) * grid_cell_size
y_length <- (62 + 6) * grid_cell_size  # Additional 2 grid cell sizes (1 for upper and 1 for lower)

# Calculate the coordinates of the other corners
upper_right_corner <- c(x = upper_left_corner[1] + x_length, y = upper_left_corner[2])
lower_right_corner <- c(x = upper_right_corner[1], y = upper_right_corner[2] - y_length)
lower_left_corner <- c(x = upper_left_corner[1], y = lower_right_corner[2])

# Create the coordinates for the rectangular polygon
rectangle_coords <- rbind(
  c(upper_left_corner[1], upper_left_corner[2]),
  c(upper_right_corner[1], upper_right_corner[2]),
  c(lower_right_corner[1], lower_right_corner[2]),
  c(lower_left_corner[1], lower_left_corner[2]),
  c(upper_left_corner[1], upper_left_corner[2])
)

# Create the sf object
rectangle_mah <- st_polygon(list(rectangle_coords)) %>%
  st_sfc(crs = st_crs(mahama_boundary))

# Plot the rectangle_mah and mahama_boundary
mapview(rectangle_mah) + mapview(mahama_boundary)

```
# Creating a grid (regular resselation)

```{r Create Grid}
grid <- rectangle_mah %>%
  st_make_grid(cellsize = c(grid_cell_size, grid_cell_size, square = T)) %>%
  #st_intersection(mahama_boundary) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(id = row_number()) %>%
  mutate(grid_area = st_area(.)) # Calculate grid cell area 

plot(grid$geometry)
plot(mah_buildings, add = TRUE, col = "red")

# Print the grid cell size
print(paste("Grid cell size:", grid_cell_size, "meters"))

# buildings_grid <- grid %>%
#   st_join(mah_buildings) %>%
#   group_by(id) %>%
#   summarize(num_buildings = n())
# 
# plot(buildings_grid["num_buildings"])

mapview(grid)
```

# ----------------------------

## Calculation indicator layer

# Population density 

Based on UNHCR data, there are 6,928 family shelters (a shelter is designed to accommodate two families) in Mahama.

The camp population in the Mahama refugee camp is at a number of 62486 individual as of September 2023. To distribute the total population evenly across buildings, taking into account the area of each building so that larger buildings can house more people.

1. Calculation of total area of all buildings 
2. Calculate the overall population density, which is the total population divided by the total area of all buildings.
3. Multiply the area of each building by the population density to estimate the number of people each building can house.
```{r Average Camp Population}
# Camp population (Sept. 2023)
camp_pop_23 <- 62486

# Number of buildings 
n_buildings_mah <- nrow(mah_buildings)
print("# Building Mahama Camp")
n_buildings_mah

# Number of shelter
n_shelter_mah <- nrow(mah_shelter)
print("# Shelter Mahama Camp")
n_shelter_mah

# Calculate average population per shelter
average_population <- camp_pop_23 / nrow(mah_shelter)
print("Average number of people per shelter")
average_population

```
The average occupancy of a shelter is almost 8 people per shelter. This number seems to be in line with the UNHCR data, which states that a shelter is designed to accommodate two families. 

```{r Distribute population count}
# Assign average occupancy number to each shelter to get the spatial distribution of the camp population 
mah_shelter$estimated_people <- average_population

# Now, the 'shelter' data frame should contain the estimated number of people each building can house

## OPEN QUESTION: should the number of people be scaled to the shelter size??
```

```{r Shelter count}
# Calculate centroid for each shelter
mah_shelter_centroids <- st_centroid(mah_shelter)

# Assign shelters to grid cells based on centroid
shelter_grid_join <- st_join(mah_shelter_centroids, grid, join = st_intersects)

# Count number of shelters per grid cell
shelter_count_per_grid <- shelter_grid_join %>%
  group_by(id) %>%
  summarize(shelter_count = n()) %>%
  ungroup()

class(shelter_count_per_grid)

# Join the shelter count with the grid
grid <- st_join(grid, shelter_count_per_grid, join = st_intersects, left = TRUE)
grid$shelter_count <- replace(grid$shelter_count, is.na(grid$shelter_count), 0) # replace NA values with 0 

# Drop the 'id.y' column
grid <- grid[, !grepl("id.y", names(grid))]
# Rename id.x to id
names(grid)[names(grid) == "id.x"] <- "id"

# Create the map of Shelter Counts per Grid
tmap_mode("plot")
tm_shape(grid) +
  tm_polygons("shelter_count", palette = "viridis", title = "# shelter") +
tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)
```

```{r Plot Population Count}
# Plot the grid
plot(grid$geometry)

# Plot buildings on top of the grid
plot(mah_shelter, add = TRUE, col = "red")

# Join shelter with grid cells and summarize the number of people in each grid cell
population_grid <- grid %>%
  st_join(mah_shelter) %>%
  group_by(id) %>%
  summarize(
    estimated_people = sum(!is.na(id)),
    shelter_count = n_distinct(id))

# Transform estimated_people == 1 to 0
population_grid <- population_grid %>%
  mutate(estimated_people = ifelse(estimated_people == 1, 0, estimated_people))

# Plot the number of buildings per grid cell 
plot(population_grid["estimated_people"])

# Plot the tmap with custom breaks for color classification
tmap_mode("plot")
tm_shape(grid) +
  tm_polygons() +
  tm_shape(population_grid) +
  tm_fill("estimated_people", 
          title = "# people", 
          style = "pretty",
          breaks = c(0, 1, 7, 13, 19),  # Custom breaks for color classification
          labels = c("0", "1-7", "7-13", "13-19"),  # Labels for the classes
          palette = "viridis") +  # Color palette 
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)

  # tm_layout(title = "Estimated Count per Grid Cell",
  #           title.position = c("center", "top"))  # Set the position of the title
```


```{r Population density}

# Calculate population density
population_grid$population_density <- population_grid$estimated_people / (30 * 30)  # population is counted for a 30x30 meter grid cell

# Normalize population density
min_density <- min(population_grid$population_density)
max_density <- max(population_grid$population_density)
population_grid$normalized_density <- (population_grid$population_density - min_density) / (max_density - min_density)

# Now 'population_grid' contains the population density and normalized density

# Plot the tmap with normalized density: Normalized Population Density per Grid Cell
tm_shape(grid) +
  tm_polygons() +
  tm_shape(population_grid) +
  tm_fill("normalized_density", 
          title = "norm. pop. density", 
          palette = "viridis",
          breaks = c(0, 0.25, 0.50, 0.75, 1),
          labels = c("0 - 0.25", "0.250 - 0.50", "0.50 - 0.75", "0.75 - 1")) +  
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)
```

# Vulnerable Groups

```{r Existence of vulnerable groups}
# Join polygons with grid cells
vul_groups_grid <- grid %>%
  st_join(mah_vul_groups) 

# Transform existence of vulnerable groups == 1, NA == 0
vul_groups_grid <- vul_groups_grid %>%
  mutate(vul_group = ifelse(!is.na(Class), 1, 0))

# Plot the tmap of Existence of vulnerable groups
tmap_mode("plot")
tm_shape(grid) +
  tm_polygons() +
  tm_shape(vul_groups_grid) +
  tm_fill("vul_group", 
          title = "Vulnerable Groups", 
          n = 2,
          palette = "viridis",
          labels = c("No Vulnerable Groups", "Vulnerable Groups")) +  
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)

```


# Critical Infrastructure 

Existence of fragile infrastructure + Sanitation Network + Drainage System

```{r Existence of fragile CI}
# Join polygons with grid cells
ci_grid <- grid %>%
  st_join(mah_ci) 

# Transform existence of critical infrastructure == 1, NA == 0
ci_grid <- ci_grid %>%
  mutate(ci = ifelse(!is.na(Type), 1, 0))

# Plot the tmap of Existence of vulnerable groups
tmap_mode("plot")
tm_shape(grid) +
  tm_polygons() +
  tm_shape(ci_grid) +
  tm_fill("ci", 
          title = "Critical Infrastructure", 
          n = 2,
          palette = "viridis",
          labels = c("No critical infrastructure", "Critical infrastructure")) +  
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)
```

```{r Sanitary Buildings count}
# Calculate centroid for each sanitary unit
mah_sanitary_centroids <- st_centroid(mah_sanitary)

# Assign sanitarys to grid cells based on centroid
sanitary_grid_join <- st_join(mah_sanitary_centroids, grid, join = st_intersects)

# Count number of sanitarys per grid cell
sanitary_count_per_grid <- sanitary_grid_join %>%
  group_by(id) %>%
  summarize(sanitary_count = n()) %>%
  ungroup()

class(sanitary_count_per_grid)

# Join the sanitary count with the grid
sanitary_grid <- st_join(grid, sanitary_count_per_grid, join = st_intersects, left = TRUE)
sanitary_grid$sanitary_count <- replace(sanitary_grid$sanitary_count, is.na(sanitary_grid$sanitary_count), 0) # replace NA values with 0 

# Drop the 'id.y' column
sanitary_grid <- sanitary_grid[, !grepl("id.y", names(sanitary_grid))]
# Rename id.x to id
names(sanitary_grid)[names(sanitary_grid) == "id.x"] <- "id"

# Create the map of sanitary Counts per Grid
tmap_mode("plot")
tm_shape(sanitary_grid) +
  tm_polygons("sanitary_count", palette = "viridis", title = "# sanitary") +
tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)
```

```{r Calculate normalized sanitary density }
# Normalize sanitary count per grid cell

sanitary_density_grid <- sanitary_grid

# Calculate sanitary density
sanitary_density_grid$sanitary_density <- sanitary_density_grid$sanitary_count / (30 * 30)  # sanitarys are counted for a 30x30 meter grid cell

# Normalize sanitary density
min_density_s <- min(sanitary_density_grid$sanitary_density)
max_density_s <- max(sanitary_density_grid$sanitary_density)
sanitary_density_grid$normalized_density <- scale(sanitary_density_grid$sanitary_density, 
                                                 center = min_density_s, 
                                                 scale = max_density_s - min_density_s)

sanitary_density_grid <-  sanitary_density_grid %>% 
  mutate(normalized_density = as.numeric(normalized_density))  # Convert to numeric
  
# Plot sanitary density
tm_shape(grid) +
  tm_polygons() +
  tm_shape(sanitary_density_grid) +
  tm_fill("normalized_density", 
          title = "Sanitary Units Density", 
          palette = "viridis", 
          breaks = c(0, 0.01, 0.33, 0.66, 1),  # Adjust breaks to have 4 classes
          na.value = 0) +  # Set NA values to 0
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)

mapview(sanitary_density_grid, zcol = "normalized_density", layer.name = "Sanitary Density") + 
  mapview(mah_sanitary, layer.name = "sanitarys", alpha.regions = 0.5, legend = FALSE, add = TRUE)
```

Drainage System
```{r Drainage Grid}
# Join drainage polygon with grid 
drainage_grid <- grid %>%
  st_join(mah_drainage)

# Transform existence of drainage systems == 1, NA == 0
drainage_grid <- drainage_grid %>%
  mutate(drainage = ifelse(!is.na(Type), 1, 0)) %>%
  mutate(rank = ifelse(!is.na(Type), 2, 0)) # assign vulnerability rank 

# Plot the tmap of Existence of vulnerable groups
tmap_mode("plot")
tm_shape(grid) +
  tm_polygons() +
  tm_shape(drainage_grid) +
  tm_fill("drainage", 
          title = "Drainage System", 
          n = 2,
          palette = "viridis",
          labels = c("No critical infrastructure", "Critical infrastructure")) +  
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)
```

# Facilities

Facility vulnerability ranks
```{r Facility ranks}
# Extract unique class names
class_names <- unique(facility_buildings$Facility)

# Output the unique class names
print(class_names)

# Create a dataframe with class names and their ranks for social and physical vulnerability
facility_classes <- c("School", "Health Center", "Cultural facilities", "Youth/women centers", "Administrative Facilities", "Security", "Nutrition Center", "Distribution center", "Open Spaces")
social_vulnerability_ranks <- c(2, 3, 2, 2, 2, 3, 3, 2, 1)  # Assign ranks for social vulnerability
physical_vulnerability_ranks <- c(2, 2, 3, 2, 2, 2, 2, 2, 0)  # Assign ranks for physical vulnerability

# Create a dataframe
facility_vulnerability <- data.frame(Facility = facility_classes, 
                                     Social_Vulnerability_Rank = social_vulnerability_ranks,
                                     Physical_Vulnerability_Rank = physical_vulnerability_ranks)

# Print the dataframe
print(facility_vulnerability)

# Join the dataframes by the "Facility" column
facility_buildings_ranked <- left_join(facility_buildings, facility_vulnerability, by = "Facility") 
print(facility_buildings_ranked)

# Remove columns 6 through 13 and 15 through 17
#facility_buildings_ranked <- facility_buildings_ranked[, -c(6:13, 15:17)]

mapview(facility_buildings_ranked, zcol = "Physical_Vulnerability_Rank")
```

First, calculate number of facilities per grid cell, meaning that when a building intersects a grid cell > add 1 to the count. 

To calculate the facilities social vulnerability of each grid cell based on the normalized area of facility buildings and their respective weights, the following steps where taken:

```{r Facilities social vulnerability per grid cell}

# Step 1: Intersect facility buildings with grid cells
facility_grid_intersect <- st_intersection(facility_buildings_ranked, grid)

# Check by plotting:
# mapview(facility_grid_intersect)

# Step 2: Calculate the area of each intersection
facility_grid_intersect <- facility_grid_intersect %>%
  mutate(intersection_area = st_area(geometry))

# Step 3: Group by facility class and grid cell ID
facility_grid_grouped <- facility_grid_intersect %>%
  group_by(Facility, id, grid_area) %>%
  summarize(total_area = sum(intersection_area)) # per facility class

# Step 4: Calculate the relative area by normalizing total area by grid area

# Convert the units of total_area to match the units of grid_area
facility_grid_grouped <- facility_grid_grouped %>%
  mutate(total_area = as.numeric(total_area))  # Convert total_area to numeric

# Now calculate the relative area by normalizing total area by grid area
facility_grid_grouped <- facility_grid_grouped %>%
  mutate(relative_area = total_area / grid_area) %>%
  ungroup()  # Remove grouping

# Weighted values 
# Merge class ranks with facility_grid_grouped based on class names
facility_grid_grouped <- merge(facility_grid_grouped, facility_vulnerability, by.x = "Facility", by.y = "Facility", all.x = TRUE)

# Calculate weighted sum of relative areas using provided ranks

#SOC
facility_grid_grouped_soc <- facility_grid_grouped %>%
  mutate(weighted_sum_soc = relative_area * (social_vulnerability_ranks))

# PHY
facility_grid_grouped_phy <- facility_grid_grouped %>%
  mutate(weighted_sum_phy = relative_area * (physical_vulnerability_ranks))
```

```{r Check for invalid geometries}
#SOC

# Check for and potentially fix invalid geometries
facility_grid_grouped_soc_valid <- facility_grid_grouped_soc %>%
 mutate(valid_geometry = st_is_valid(geometry))

# Check for invalid geometries
invalid_geometries <- facility_grid_grouped_soc_valid %>%
  filter(!valid_geometry)

# Print output if there are invalid geometries
if(nrow(invalid_geometries) > 0) {
  print("Invalid geometries found:")
  print(invalid_geometries)
} else {
  print("No invalid geometries found.")
}

# If there are invalid geometries, you can attempt to fix them
facility_grid_grouped_soc_fixed <- facility_grid_grouped_soc_valid %>%
  filter(valid_geometry == FALSE) %>%
  mutate(geometry = st_make_valid(geometry))

# Now recheck for validity
facility_grid_grouped_soc_fixed <- facility_grid_grouped_soc_fixed %>%
  mutate(valid_geometry = st_is_valid(geometry))

# Combine the fixed geometries with the rest of the data
facility_grid_grouped_soc <- rbind(facility_grid_grouped_soc_valid %>% filter(valid_geometry == TRUE), facility_grid_grouped_soc_fixed)

# PHY

# Check for and potentially fix invalid geometries
facility_grid_grouped_phy_valid <- facility_grid_grouped_phy %>%
  mutate(valid_geometry = st_is_valid(geometry))

# Check for invalid geometries
invalid_geometries <- facility_grid_grouped_phy_valid %>%
  filter(!valid_geometry)

# Print output if there are invalid geometries
if(nrow(invalid_geometries) > 0) {
  print("Invalid geometries found:")
  print(invalid_geometries)
} else {
  print("No invalid geometries found.")
}

# If there are invalid geometries:
facility_grid_grouped_phy_fixed <- facility_grid_grouped_phy_valid %>%
  filter(valid_geometry == FALSE) %>%
  mutate(geometry = st_make_valid(geometry))

# Now recheck for validity
facility_grid_grouped_phy_fixed <- facility_grid_grouped_phy_fixed %>%
  mutate(valid_geometry = st_is_valid(geometry))

# Combine the fixed geometries with the rest of the data
facility_grid_grouped_phy <- rbind(facility_grid_grouped_phy_valid %>% filter(valid_geometry == TRUE),
                               facility_grid_grouped_phy_fixed)

```

```{r Summarize and Normalize grid counts}
# SOC

#----
# Combine the (fixed) geometries with the rest of the data
#facility_grid_grouped_soc <- rbind(facility_grid_grouped_soc %>% filter(valid_geometry == TRUE),
#                               facility_grid_grouped_soc_fixed)

# Summarize the data per grid cell
facility_grid_summed_soc <- facility_grid_grouped_soc %>%
  filter(id != 120) %>% # since this geometry is causing trouble but very small so it can be neglected
  group_by(id) %>%
  summarize(sum_weighted_values = sum(weighted_sum_soc, na.rm = TRUE)) 


# Normalize the summed values to be within a score of 0 and 1
max_sum <- max(facility_grid_summed_soc$sum_weighted_values)
min_sum <- min(facility_grid_summed_soc$sum_weighted_values)

facility_grid_grouped_soc <- facility_grid_summed_soc %>%
  mutate(normalized_score = (sum_weighted_values - min_sum) / (max_sum - min_sum))

sf_use_s2(FALSE)

#PHY

# Combine the (fixed) geometries with the rest of the data
#facility_grid_grouped_phy <- rbind(facility_grid_grouped_phy %>% filter(valid_geometry == TRUE),
#                               facility_grid_grouped_phy_fixed)

# Summarize the data
facility_grid_summed_phy <- facility_grid_grouped_phy %>%
  group_by(id) %>%
  summarize(sum_weighted_values = sum(weighted_sum_phy, na.rm = TRUE)) %>%
  filter(id != 120) # since this geometry is causing trouble but very small so it can be neglected

# Normalize the summed values to be within a score of 0 and 1
max_sum <- max(facility_grid_summed_phy$sum_weighted_values)
min_sum <- min(facility_grid_summed_phy$sum_weighted_values)

facility_grid_grouped_phy <- facility_grid_summed_phy %>%
  mutate(normalized_score = (sum_weighted_values - min_sum) / (max_sum - min_sum))
```

```{r Clead facility df}
#SOC
# Drop the geometry column from the sf object for later join
facility_grid_grouped_soc_df <- st_drop_geometry(facility_grid_grouped_soc)

# Left join the full grid with the summarized data
facility_soc_grid <- left_join(grid, facility_grid_grouped_soc_df, by = "id")

# Convert normalized_score column to numeric
facility_soc_grid$normalized_score <- as.numeric(as.character(facility_soc_grid$normalized_score))

# Replace NA values in the normalized_score column with 0
facility_soc_grid$normalized_score[is.na(facility_soc_grid$normalized_score)] <- 0

# Tidy df
facility_soc_grid$sum_weighted_values <- NULL


#PHY
# Drop the geometry column from the sf object for later join
facility_grid_grouped_phy_df <- st_drop_geometry(facility_grid_grouped_phy)

# Left join the full grid with the summarized data
facility_phy_grid <- left_join(grid, facility_grid_grouped_phy_df, by = "id")

# Convert normalized_score column to numeric
facility_phy_grid$normalized_score <- as.numeric(as.character(facility_phy_grid$normalized_score))

# Replace NA values in the normalized_score column with 0
facility_phy_grid$normalized_score[is.na(facility_phy_grid$normalized_score)] <- 0

# Tidy df
facility_phy_grid$sum_weighted_values <- NULL

```


```{r Plotting facilities vulnerability}
# Plot the tmap
tmap_mode("plot")

# SOC
tm_shape(facility_soc_grid) +
  tm_fill("normalized_score", 
          title = "Facilities of social importance", 
          palette = "viridis", 
          breaks = c(0, 0.01, 0.33, 0.66, 1),  # Adjust breaks to have 4 classes
          na.value = 0) +  # Set NA values to 0
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)

# PHY
tm_shape(grid) +
  tm_polygons() +
  tm_shape(facility_phy_grid) +
  tm_fill("normalized_score", 
          title = "Facilities of physical importance", 
          palette = "viridis", 
          breaks = c(0, 0.01, 0.33, 0.66, 1),  # Adjust breaks to have 4 classes
          na.value = 0) +  # Set NA values to 0
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)

# tm_shape(facility_buildings) +
#   tm_polygons()

```

# Land use 

Data download from here: https://storage.googleapis.com/earthenginepartners-hansen/GLCLU2000-2020/v2/download.html 

```{r Land use}
# # Clip raster layer to boundary
# clipped_raster <- crop(GLCLUC2020, extent(mahama_boundary))
# mah_land_use_raster <- mask(clipped_raster, mahama_boundary)
# 
# class(grid)
# 
# # Plot the clipped raster to visualize
# 
# plot(mah_land_use_raster,
#      col = names,
#      main="Land Cover GLCLUC2020",
#      axes = F)
# plot(grid, add = T, col = "transparent")
# plot(mah_buildings, add = T)
# 
# # Extracted raster cells and percentage of area
# # covered within each polygon
# valuesatcell <- extract(mah_land_use_raster, grid, na.rm = TRUE, weights = TRUE)
# 
# # Change column names
# colnames(valuesatcell)[1] <- "id"
# colnames(valuesatcell)[2] <- "land_use_class"
```

```{r}
# Classification of land cover classes
# 
# land_use_classes <- unique(valuesatcell$land_use_class)
# print(land_use_classes)
# 
# # Create a dataframe with class names and their ranks for social and physical vulnerability
# lu_classes <- c(250, 244, 22, 23, 24, 205, 203, 207, 202, 20, 16, 17, 18, 124, 25, 19, 26, 206, 27, 21)
# cover_name <- c("Built-up", "Cropland", "Dense short vegetation", "Dense short vegetation", "Dense short vegetation", "Open surface water",  "Open surface water", "Open surface water", "Open surface water", "Dense short vegetation", "Semi-arid", "Semi-arid", "Semi-arid", "Dense short vegetation", "Tree cover", "Dense short vegetation", "Tree cover", "Open surface water", "Tree cover", "Dense short vegetation")  
# vulnerability_rank <- c(0, 2, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
# 
# # Create a dataframe
# lu_classes <- data.frame(class = lu_classes, 
#                                     name = cover_name, 
#                                     rank = vulnerability_rank)
# 
# # Join classes and ranks df with valuesatcell
# lu_cell <- left_join(valuesatcell, lu_classes, by = c("land_use_class" = "class"))
```


# Land cover with UNHCR data 

- agricultural areas: 3
- open spaces: 2

```{r Land Use grid data}
mapview(mah_land_cover, zcol = "Type")

mah_land_cover <- mah_land_cover %>%
  mutate(Rank = ifelse(Type == "Agricultural Land", 3,
                ifelse(Type == "Open Space", 2, NA))) %>%
  mutate(norm = scale(Rank, center = FALSE, scale = max(Rank, na.rm = TRUE))) %>%
  mutate(norm = as.numeric(norm))

# Join polygons with grid cells
cover_grid <- grid %>%
  st_join(mah_land_cover) %>%
  mutate(norm = ifelse(is.na(norm), 0, norm))

# Plot the tmap 
tmap_mode("plot")
tm_shape(grid) +
  tm_polygons() +
  tm_shape(cover_grid) +
  tm_fill("norm", 
          title = "Land Cover",
          breaks = c(0, 0.25, 0.5, 0.75, 1),  # Adjust breaks to have 4 classes
          labels = c("0", "1", "2", "3"),
          palette = "viridis") +  
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)
```

# Shelter type
```{r Plot Shelter Type}
class(shelter_count_per_grid)

# Create the map of Shelter Counts per Grid
tmap_mode("plot")

# Plot the grid polygons with shelter counts
tm_grid <- tm_shape(grid) +
  tm_polygons("shelter_count", palette = "viridis", title = "# shelter")

# Plot the shelter types on the map
tm_shelters <- tm_shape(mah_shelter_type) +
  tm_symbols(col = "Type", shape = 16, size = 0.5) +  # Adjust size and shape as needed
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)

# Combine the layers
tm_combine <- tm_grid + tm_shelters

# Plot the map
tm_combine
```
Assign higher vulnerability score to the 

Abandoned shelters location form mapping workshop: not quite sure where exactly but but structurally sensitive!! 

Shelter Physical Vulnerability: 

1. Emergency: Habitable covered living space providing a secure and healthy environment with privacy and dignity. The shelters typically simple, one room structures implemented to provide critical life saving emergency assistance
  --> vulnerability rank: 3

2. Transitional: A range of shelter options that help populations affected by a humanitarian crises progress from an initial emergency arrangement to a more suitable shelter solution, better adapted to their needs in terms of habitability.
  --> vulnerability rank: 2

3. Durable: Beyond the emergency and transitional phase, shelters that are adapted and contextualized according to the following elements: climate, cultural practice and habits, local availability of skills, access to adequate construction materials and geographical context.
  --> vulnerability rank: 1
  
4. Abandoned: Not inhabited, fragile structure
  --> vulnerability rank: 1
  
```{r Assign Shelter Type Vulnerability Ranks}
# Assign vulnerability ranks 
mah_shelter_type <- mah_shelter_type %>%
  mutate(rank = case_when(
    Type %in% c("Emergency") ~ 3,
    Type == "Transitional" ~ 2,
    Type == "Durable" ~ 1,
    Type == "Abandoned" ~ 1,
    TRUE ~ NA_integer_
  ))

# Intersect shelter type polygon with grid 
shelter_type_intersect <- st_intersection(mah_shelter_type, grid) 

# Drop the geometry column from the sf object for later join
shelter_type_intersect_df <- st_drop_geometry(shelter_type_intersect)

# Join polygons with grid cells
shelter_type_grid <- left_join(grid, shelter_type_intersect_df, by = "id")

# Transform existence of shelter type == 1, NA == 0
shelter_type_grid <- shelter_type_grid %>%
  mutate(exist = ifelse(!is.na(Type), 1, 0))

#mapview(shelter_type_grid, zcol = "exist")
```

```{r Calculate normalized shelter density }
# Normalize shelter count per grid cell

shelter_density_grid <- grid

# Calculate shelter density
shelter_density_grid$shelter_density <- shelter_density_grid$shelter_count / (30 * 30)  # shelters are counted for a 30x30 meter grid cell

# Normalize shelter density
min_density_s <- min(shelter_density_grid$shelter_density)
max_density_s <- max(shelter_density_grid$shelter_density)
shelter_density_grid$normalized_density <- scale(shelter_density_grid$shelter_density, 
                                                 center = min_density_s, 
                                                 scale = max_density_s - min_density_s)

shelter_density_grid <-  shelter_density_grid %>% mutate(normalized_density = as.numeric(normalized_density))  # Convert to numeric

# Plot shelter density
tm_shape(grid) +
  tm_polygons() +
  tm_shape(shelter_density_grid) +
  tm_fill("normalized_density", 
          title = "Shelter Density", 
          palette = "viridis", 
          breaks = c(0, 0.01, 0.33, 0.66, 1),  # Adjust breaks to have 4 classes
          na.value = 0) +  # Set NA values to 0
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)

mapview(shelter_density_grid, zcol = "normalized_density", layer.name = "Shelter Density") + 
  mapview(mah_shelter, layer.name = "Shelters", alpha.regions = 0.5, legend = FALSE, add = TRUE)
```
2.	Join shelter type grid with the shelter density grid by grid cell id
3.	if shelter type “exist” == 1 then multiply the shelter density “normalized_density” by the rank “rank” and store the value in a new column “weight”, if “exist” == 0, then just take “normalized_density” and store it in weight
4.	normalize the weight [0;1]

```{r Plot Shelter Type Density}
# Drop the geometry column from the sf object for later join
shelter_density_grid_df <- st_drop_geometry(shelter_density_grid)

# Step 2: Join shelter type grid with the shelter density grid by grid cell id
shelter_density_type_grid <- left_join(shelter_type_grid, shelter_density_grid_df, by = "id")

# Step 2: Create the weight column based on conditions
shelter_density_type_grid <- shelter_density_type_grid %>%
  mutate(weight = ifelse(exist == 1, normalized_density * rank, normalized_density))

# Step 3: Normalize the weight between [0,1]
min_weight <- min(shelter_density_type_grid$weight, na.rm = TRUE)
max_weight <- max(shelter_density_type_grid$weight, na.rm = TRUE)
shelter_density_type_grid$normalized_w <- scale(shelter_density_type_grid$weight, 
                                                center = min_weight,
                                                scale = max_weight - min_weight)

shelter_density_type_grid <-  shelter_density_type_grid %>% mutate(normalized_w = as.numeric(normalized_w))  # Convert to numeric

# Plot the final result
tmap_mode("plot")
tm_shape(shelter_density_type_grid) +
  tm_polygons("normalized_w", 
              title = "Normalized Weight", 
              palette = "viridis", 
              breaks = c(0, 0.01, 0.33, 0.66, 1),  # Adjust breaks to have 4 classes
              na.value = 0) +  # Set NA values to 0
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)
# tm_shape(mah_shelter_type)+
#   tm_polygons() +
# tm_shape(mah_shelter) +
#   tm_polygons()

```

# Roads / Transport

Rank the road types 
```{r Road ranks}
# Rename fcalss column
mah_osm_roads <- mah_osm_roads %>% rename("Type" = "fclass")
head(mah_osm_roads)

# Extract unique road class names
road_types <- unique(mah_osm_roads$Type)
print(road_types)

mapview(mah_osm_roads, zcol = "Type")

# Create a dataframe with class names and their ranks
road_classes <- c("residential", "service", "unclassified", "path", "footway")
road_ranks <- c(2, 3, 1, 1, 1)  # Assign ranks for social vulnerability

# Create a dataframe
road_vulnerability <- data.frame(Type = road_classes, 
                                     Rank = road_ranks)

# Print the dataframe
print(road_vulnerability)

# Join the dataframes by the "Facility" column
roads_ranked <- left_join(mah_osm_roads, road_vulnerability, by = "Type") 
head(roads_ranked)

#mapview(grid, col.regions = "transparent") + mapview(roads_ranked, zcol = "Rank")

```

```{r Roads Vector data}
tmap_mode("plot")
```

```{r Roads Grid Intersect}
# Intersect road type polygon with grid 
road_type_intersect <- st_intersection(roads_ranked, grid) 
head(road_type_intersect)

# Drop the geometry column from the sf object for later join
road_type_intersect_df <- st_drop_geometry(road_type_intersect)

# Summarize the data
road_grid_summed <- road_type_intersect_df %>%
  group_by(id) %>%
  summarize(sum_ranks = sum(Rank, na.rm = TRUE))

# Normalize the summed values to be within a score of 0 and 1
max_sum_r <- max(road_grid_summed$sum_ranks)
min_sum_r <- min(road_grid_summed$sum_ranks)

road_grid_summed <- road_grid_summed %>%
  mutate(normalized_score = (sum_ranks - min_sum_r) / (max_sum_r - min_sum_r))

# Convert column to numeric 
road_grid_summed <- road_grid_summed %>% mutate(normalized_score = as.numeric(normalized_score))  # Convert to numeric

# Join polygons with grid cells
road_type_grid <- left_join(grid, road_grid_summed, by = "id")

# Transform existence of road segments within an grid to 1, NA == 0
road_type_grid <- road_type_grid %>%
  mutate(normalized_score = ifelse(is.na(normalized_score), 0, normalized_score)) # Replace NA values in Rank with 0 > no roads within a grid cell
```

```{r Visualization of road grid}
# Map the road grid 

# Plot the final result
tmap_mode("view")
tm_shape(road_type_grid) +
  tm_polygons("normalized_score", 
              title = "Normalized roads rank", 
              palette = "viridis", 
              breaks = c(0, 0.01, 0.33, 0.66, 1),  # Adjust breaks to have 4 classes
              na.value = 0) +  # Set NA values to 0
  tm_layout(title = "") +
  tm_borders(lwd = 0.1, col = "white") +
  tm_shape(mahama_boundary) +
  tm_borders(col = "white", lwd = 0.5) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "", 
              title.position = c("left", "bottom"), 
              frame = F)
```

# Rasterization of grid layer

```{r Resample grid}
# Create template stars object for rasterization
grid_stars <- st_as_stars(st_bbox(grid))
class(grid_stars)

grid_matrix <- grid  # Replace your_grid_data with your actual grid data

# Get the dimensions of the grid matrix
nx <- 55
ny <- 62

grid_temp <- st_rasterize(grid, st_as_stars(st_bbox(grid), nx = nx, ny = ny, values = NA_real_), align = T)

# Rasterize population_grid using the raster template
soc1_raster <- st_rasterize(population_grid %>% dplyr::select(normalized_density, geometry),
                             template = grid_temp, align = T)
```

```{r Camp extent}
# Starting with rasterizing the Mahama Camp boundary 

# Clip the raster to the extent of the polygon
bound_rast <- raster(mahama_boundary) 
class(bound_rast)
```


```{r Rasterization of indicator layer}

# Rasterize based on geometry and a column value.
soc1_raster <- st_rasterize(population_grid %>% dplyr::select(normalized_density, geometry),
                             template = grid_temp, align = T)
# Convert stars object to raster
soc1_raster <- as(soc1_raster, "Raster") %>% 
  mask(., mahama_boundary)

soc2_raster <- st_rasterize(vul_groups_grid %>% dplyr::select(vul_group, geometry),
                             template = grid_temp, align = T)
soc2_raster <- as(soc2_raster, "Raster") %>% 
  mask(., mahama_boundary)

soc3_raster <- st_rasterize(facility_soc_grid %>% dplyr::select(normalized_score, geometry),
                             template = grid_temp, align = T)
soc3_raster <- as(soc3_raster, "Raster") %>% 
  mask(., mahama_boundary)

soc4_raster <- st_rasterize(cover_grid %>% dplyr::select(norm, geometry),
                             template = grid_temp, align = T) 
soc4_raster <- as(soc4_raster, "Raster") %>% 
  mask(., mahama_boundary)

phy1_raster <- st_rasterize(shelter_density_type_grid %>% dplyr::select(normalized_w, geometry), 
                             template = grid_temp, align = T)
phy1_raster <- as(phy1_raster, "Raster") %>% 
  mask(., mahama_boundary)

phy3_raster <- st_rasterize(facility_phy_grid %>% dplyr::select(normalized_score, geometry),
                             template = grid_temp, align = T)
phy3_raster <- as(phy3_raster, "Raster") %>% 
  mask(., mahama_boundary)

phy4_raster <- st_rasterize(road_type_grid %>% dplyr::select(normalized_score, geometry),
                             template = grid_temp, align = T)
phy4_raster <- as(phy4_raster, "Raster") %>% 
  mask(., mahama_boundary)
```


*Critical Infrastructure raster*
Overlay rasters and weight by rank: 

-	Fragile shelter infrastructure -> 3
-	Sanitation Network / Sanitary units -> 3
-	Drainage System -> 2

```{r Rasterization of CI}
ci_raster <- st_rasterize(ci_grid %>% dplyr::select(ci, geometry), 
                             template = grid_temp, align = T)
ci_raster <- as(ci_raster, "Raster") %>% 
  mask(., mahama_boundary)

sanitary_raster <- st_rasterize(sanitary_density_grid %>% dplyr::select(normalized_density, geometry),
                              template = grid_temp, align = T)

sanitary_raster <- as(sanitary_raster, "Raster") %>% 
  mask(., mahama_boundary)

drainage_raster <- st_rasterize(drainage_grid %>% dplyr::select(drainage, geometry),
                              template = grid_temp, align = T)
drainage_raster <- as(drainage_raster, "Raster") %>% 
  mask(., mahama_boundary)

# Create final raster using weigthed overlay and assigned ranks 
ci_final_raster <- (ci_raster * 3) + (sanitary_raster * 3) + (drainage_raster * 2) 

# Rescale the values of ci_final_raster to range from 0 to 1
min_value <- minValue(ci_final_raster)
max_value <- maxValue(ci_final_raster)

# Rescale the values of ci_final_raster to range from 0 to 1
phy2_raster <- (ci_final_raster - min_value) / (max_value - min_value)
```


```{r Plotting individual vulnerability indicator raster maps}
# Plot each raster
plot1 <- plot(soc1_raster, main="Population Density", axes=F, box=F)
plot2 <- plot(soc2_raster, main="Vulnerability Groups", axes=F, box=F)
plot3 <- plot(soc3_raster, main="Facility SOC", axes=F, box=F)
plot4 <- plot(soc4_raster, main="Cover Grid", axes=F, box=F)

plot5 <- plot(phy1_raster, main="Shelter Density Type", axes=F, box=F)
plot6 <- plot(phy2_raster, main="CI Grid", axes=F, box=F)
plot7 <- plot(phy3_raster, main="Facility PHY", axes=F, box=F)
plot8 <- plot(phy4_raster, main="Road Type Grid", axes=F, box=F)
```

```{r Visualize individual indicator raster layer}
tmap_mode("plot")

# Define tmaps without legends
tm_soc1 <- tm_shape(soc1_raster) +
  tm_raster(palette = "YlOrRd", style = "cont") +
  tm_layout(legend.show = FALSE, title = "", title.position = c("left", "bottom"), title.size = 1, title.color = "black", frame = FALSE) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_soc2 <- tm_shape(soc2_raster) +
  tm_raster(palette = "YlOrRd",  style = "cont") +
  tm_layout(legend.show = FALSE, title = "", title.position = c("left", "bottom"), title.size = 1, title.color = "black", frame = FALSE) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_soc3 <- tm_shape(soc3_raster) +
  tm_raster(palette = "YlOrRd",  style = "cont") +
  tm_layout(legend.show = FALSE, title = "", title.position = c("left", "bottom"), title.size = 1, title.color = "black", frame = FALSE) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_soc4 <- tm_shape(soc4_raster) +
  tm_raster(palette = "YlOrRd",  style = "cont") +
  tm_layout(legend.show = FALSE, title = "", title.position = c("left", "bottom"), title.size = 1, title.color = "black", frame = FALSE) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_phy1 <- tm_shape(phy1_raster) +
  tm_raster(palette = "YlOrRd",  style = "cont") +
  tm_layout(legend.show = FALSE, title = "", title.position = c("left", "bottom"), title.size = 1, title.color = "black", frame = FALSE) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_phy2 <- tm_shape(phy2_raster) +
  tm_raster(palette = "YlOrRd",  style = "cont") +
  tm_layout(legend.show = FALSE, title = "", title.position = c("left", "bottom"), title.size = 1, title.color = "black", frame = FALSE) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_phy3 <- tm_shape(phy3_raster) +
  tm_raster(palette = "YlOrRd",  style = "cont") +
  tm_layout(legend.show = FALSE, title = "", title.position = c("left", "bottom"), title.size = 1, title.color = "black", frame = FALSE) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_phy4 <- tm_shape(phy4_raster) +
  tm_raster(palette = "YlOrRd",  style = "cont") +
  tm_layout(legend.show = F, title = "", title.position = c("left", "bottom"), title.size = 1, title.color = "black", frame = FALSE) +
  tm_legend(outside = TRUE) + 
  tm_shape(mahama_boundary) +
  tm_borders()

# Arrange tmaps
tmap_indicators <- tmap_arrange(tm_soc1, tm_soc2, tm_soc3, tm_soc4, tm_phy1, tm_phy2, tm_phy3, tm_phy4, nrow = 2)
tmap_indicators

# Save tmap 
tmap_save(tmap_indicators, "figs/indicators.png", height = 5)

png(filename = "figs/indicators_t.png", width = 5, height = 5, units = 'in', res = 300, bg = "transparent")

# Print the map to the device
print(tmap_indicators)

# Turn off the device
dev.off()
```

# Weighted raster overlay

By multiplying the individual vulnerability indicators (raster layers [0;1]) with their assigned weight (throught the AHP process) and building the sum, we get the FVI for AHP and FAHP (see Ganji et al. 2022). Weighted linear combination was used. 

indicator	wMin	wModal	wMax	Defuzzified_w	Norm_Defuzzified_w
SOC_1	0.2740	0.4033	0.6029	0.4267	0.3971
SOC_2	0.2234	0.3391	0.5249	0.3625	0.3373
SOC_3	0.1170	0.1836	0.3011	0.2005	0.1866
SOC_4	0.0443	0.0737	0.1368	0.0849	0.0790

```{r Weighted raster overlay}
# Fuzzy AHP weights
f_weight_soc1 <- 0.3971
f_weight_soc2 <- 0.3373
f_weight_soc3 <- 0.1866
f_weight_soc4 <- 0.0790

f_weight_phy1 <- 0.2962
f_weight_phy2 <- 0.2584
f_weight_phy3 <- 0.3062
f_weight_phy4 <- 0.1393

# AHP weights

weight_soc1 <- 0.3416521
weight_soc2 <- 0.3404284
weight_soc3 <- 0.2256971
weight_soc4 <- 0.0922223

weight_phy1 <- 0.2458105
weight_phy2 <- 0.2375048
weight_phy3 <- 0.3321112
weight_phy4 <- 0.1845735

```

## Calculating the HFVI 

```{r HFVI calculation}
# Weighted overlay

# overlay rasters: FAHP weights
f_overlayed_soc <- (soc1_raster * f_weight_soc1) + (soc2_raster * f_weight_soc2) + (soc3_raster * f_weight_soc3) + (soc4_raster * f_weight_soc4)
f_overlayed_phy <- (phy1_raster * f_weight_phy1) + (phy2_raster * f_weight_phy2) + (phy3_raster * f_weight_phy3) + (phy4_raster * f_weight_phy4)

# Overlay rasters: AHP weights
overlayed_soc <- (soc1_raster * weight_soc1) + (soc2_raster * weight_soc2) + (soc3_raster * weight_soc3) + (soc4_raster * weight_soc4)
overlayed_phy <- (phy1_raster * weight_phy1) + (phy2_raster * weight_phy2) + (phy3_raster * weight_phy3) + (phy4_raster * weight_phy4)
```

```{r Nomalization}
# Normalize function
normalize <- function(x) {
  (x - min(x[], na.rm = TRUE)) / (max(x[], na.rm = TRUE) - min(x[], na.rm = TRUE))
}

# Normalize the intermediate overlays
overlayed_soc <- normalize(overlayed_soc)
overlayed_phy <- normalize(overlayed_phy)
f_overlayed_soc <- normalize(f_overlayed_soc)
f_overlayed_phy <- normalize(f_overlayed_phy)
```

```{r Mapping weighted raster overlay}

# Fuzzy AHP
tm_f_soc <- tm_shape(f_overlayed_soc) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "FAHP", 
            breaks = breaks) + 
  tm_layout(legend.show = T, 
            legend.position = c("left", "bottom"),
            title = "SOC", 
            title.position = c("left", "bottom"), 
            frame = FALSE) +
    tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "",
              frame = F) + 
  tm_shape(mahama_boundary) +
  tm_borders()


tm_f_phy <- tm_shape(f_overlayed_phy) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "FAHP", 
            breaks = breaks) +
  tm_layout(legend.show = T,
            legend.position = c("left", "bottom"),
            title = "PHY", 
            title.position = c("left", "bottom"), 
            frame = FALSE) +
  tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "",
              title.position = c("left", "bottom"),
              frame = F) + 
  tm_shape(mahama_boundary) +
  tm_borders()

# Arrange tmaps
tmap_arrange(tm_f_soc, tm_f_phy, nrow = 1)

# AHP
# Fuzzy 
tm_soc <- tm_shape(overlayed_soc) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "AHP") +
  tm_layout(legend.show = T, 
            legend.position = c("left", "bottom"),
            title = "SOC", 
            title.position = c("left", "bottom"), 
            frame = FALSE) +
  tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "",
              title.position = c("left", "bottom"),
              frame = F) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_phy <- tm_shape(overlayed_phy) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "AHP") +
  tm_layout(legend.show = T,
            legend.position = c("left", "bottom"),
            title = "PHY", 
            title.position = c("left", "bottom"), 
            frame = FALSE) +
  tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "",
              title.position = c("left", "bottom"),
              frame = F) + 
  tm_shape(mahama_boundary) +
  tm_borders()

# Arrange tmaps
tmap_arrange(tm_soc, tm_phy, nrow = 1)
```

```{r Save figures}
# Arrange tmaps
dim_map <- tmap_arrange(tm_f_soc, tm_f_phy, nrow = 1)

# Save tmap 
tmap_save(dim_map, "figs/dim_map.png", width = 10, height = 5)
```

### LOW PASS FILTER for smoothing 
Using moving window calculations / focal filtering: The function is applied on the neighborhood using mean for a low-pass filter 
The purpose of the mean, or low-pass, filter is to produce a smoothed image, where extreme values (possibly noise) are cancelled out.

the lowpass matrix specifies the weights or the kernel used for the focal operation. I am using a 3x3 kernel matrix where all elements are 1. This means that the focal operation will calculate the mean (fun) within a 3x3 neighborhood around each pixel in the raster layer.

```{r Smoothed SOC and PHY vulnerability maps}

# Smoothing of the final raster layer

# Using low-pass filter
# filter=matrix(1/9, nrow=3, ncol=3) would be equivalent to mean with ngb=3 in the focal function
lowpass <- matrix(1,3,3) # 3x3 moving window / kernel

f_overlayed_soc_smoothed <- focal(f_overlayed_soc, w = lowpass, fun=mean, na.rm = TRUE) %>%
  mask(., mahama_boundary)

f_overlayed_phy_smoothed <- focal(f_overlayed_phy, w = lowpass, fun=mean, na.rm = TRUE) %>%
  mask(., mahama_boundary)

tm_soc_smoothed <- tm_shape(f_overlayed_soc_smoothed) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "SOC",
             breaks = seq(0, 1, by = 0.25)) +
  tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "",
              title.position = c("left", "bottom"),
              frame = F) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tm_phy_smoothed <- tm_shape(f_overlayed_phy_smoothed) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "PHY",
            breaks = seq(0, 1, by = 0.25)) + 
  tm_layout(legend.show = T,
              legend.title.color = "black",
              legend.position = c("left", "bottom"),
              legend.text.color = "black",
              title = "",
              title.position = c("left", "bottom"),
              frame = F) + 
  tm_shape(mahama_boundary) +
  tm_borders()

tmap_arrange(tm_soc_smoothed, tm_phy_smoothed, nrow = 1)

```

```{r Resacle domain ranks}

# AHP
soc_w <- 0.1840118
phy_w <- 0.2583137

# Calculate the sum
max_value <- soc_w + phy_w

# Normalize the values
scaled_value_soc <- soc_w / max_value
scaled_value_phy <- phy_w / max_value

# Print the results
scaled_value_soc
scaled_value_phy


# FAHP importance of SOC vs PHY

soc_w_f <- 0.2194
phy_w_f <- 0.2345

# Calculate the sum
max_value_f <- soc_w_f + phy_w_f

# Normalize the values
scaled_value_soc_f <- soc_w_f / max_value_f
scaled_value_phy_f <- phy_w_f / max_value_f

# Print the results
scaled_value_soc_f
scaled_value_phy_f

```

```{r Total overlay}
# AHP
tot_vul <- overlayed_soc * scaled_value_soc + overlayed_phy * scaled_value_phy
# Convert stars object to raster
tot_vul_raster <- as(tot_vul, "Raster")

#FAHP
tot_f_vul <- f_overlayed_soc * scaled_value_soc_f + f_overlayed_phy * scaled_value_phy_f

# Convert stars object to raster
tot_f_vul_raster <- as(tot_f_vul, "Raster")
```

```{r Normalization}
# Use normalize function defined above

# Normalize the intermediate overlays
tot_vul_raster <- normalize(tot_vul_raster)
tot_f_vul_raster <- normalize(tot_f_vul_raster)
```

# Mapping the HFVI

```{r Total vulnerability visualization}

# Define breakpoints for four classes
breaks <- c(c(0, 0.01, 0.3333, 0.6666, 1))

# Define labels for the legend
labels <- c("0", "0.01 - 0.33", "0.33 - 0.66", "0.66 - 1")

# ----

tmap_mode("view")

tm_vul <- tm_shape(tot_vul_raster) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "AHP", 
            alpha = 0.7,
            breaks = seq(0, 1, by = 0.25)) +
  tm_layout(legend.show = TRUE,
            legend.position = c("left", "bottom"),
            frame = FALSE) +
  tm_shape(mahama_boundary) +
  tm_borders(col = "black", lwd = 0.7) +
  tm_layout(legend.show = TRUE,
            legend.title.color = "black",
            legend.position = c("left", "bottom"),
            legend.text.color = "black",
            title = "",
            title.position = c("left", "bottom"),
            frame = FALSE) +

  # Basemap options
  tm_basemap("Esri.WorldImagery") +  # Basemap 1
  tm_basemap("Esri.WorldTopoMa") +  # Basemap 2
  tm_basemap("OpenStreetMap.Mapnik")   # Basemap 3

tm_vul

tm_f_vul <- tm_shape(tot_f_vul_raster) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "FAHP", 
            alpha = 0.7,
            breaks = seq(0, 1, by = 0.25)) + 
  tm_layout(legend.show = TRUE,
            legend.position = c("left", "bottom"),
            frame = FALSE) +
  tm_shape(mahama_boundary) +
  tm_borders(col = "black", lwd = 0.5) +
  tm_layout(legend.show = TRUE,
            legend.title.color = "black",
            legend.position = c("left", "bottom"),
            legend.text.color = "black",
            title = "",
            title.position = c("left", "bottom"),
            frame = FALSE) +
  
    # Basemap options
  tm_basemap("Esri.WorldImagery") +  # Basemap 1
  tm_basemap("Esri.WorldTopoMa") +  # Basemap 2
  tm_basemap("OpenStreetMap.Mapnik")   # Basemap 3

tm_f_vul
```


```{r Smoothing filter}

# Smoothing of the final raster layer using the defined low-pass filter 
tot_vul_raster_smoothed <- focal(tot_vul_raster, w = lowpass, fun=mean, na.rm = TRUE) %>%
  mask(., mahama_boundary)
tot_vul_raster_smoothed <- normalize(tot_vul_raster_smoothed) # Normalize

tot_f_vul_raster_smoothed <- focal(tot_f_vul_raster, w = lowpass, fun=mean, na.rm = TRUE) %>%
  mask(., mahama_boundary)
tot_f_vul_raster_smoothed <- normalize(tot_f_vul_raster_smoothed) # Normalize

# Plotting
tmap_mode("plot")
# Set tmap options
tmap_options(legend.position = c("left", "bottom"), legend.outside = TRUE, legend.frame = FALSE, legend.title.color = "black", legend.text.color = "black", title.position = c("left", "bottom"), frame = F)

# Create the tm_plot object for FAHP
tm_vul_smoothed <- tm_shape(tot_vul_raster_smoothed) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "AHP",
            breaks = seq(0, 1, by = 0.25)) + 
  tm_shape(mahama_boundary) +
  tm_borders()

# Plot the object
tm_vul_smoothed

# Create the tm_plot object for FAHP
tm_f_vul_smoothed <- tm_shape(tot_f_vul_raster_smoothed) +
  tm_raster(palette = "YlOrRd",
            style = "cont",
            title = "FAHP",
            breaks = seq(0, 1, by = 0.25)) + 
  tm_shape(mahama_boundary) +
  tm_borders()

# Plot the object
tm_f_vul_smoothed

# Arrange tmaps
tmap_arrange(tm_vul_smoothed, tm_f_vul_smoothed, nrow = 1)
```

```{r Histrogram of index values}
hist(tot_vul_raster)
hist(tot_f_vul_raster)

# Compute the histogram of raster values
hist_values <- hist(values(tot_f_vul_raster), plot = FALSE)
hist_values

# Determine classification intervals (example: equal intervals)
num_breaks <- 4  # Specify the number of intervals
breaks <- seq(minValue(tot_f_vul_raster), maxValue(tot_f_vul_raster), length.out = num_breaks + 1)
breaks

# Define labels for each rank
labels <- c("none", "low", "medium", "high")

# Classify the raster values into different categories
classified_raster <- cut(values(tot_f_vul_raster), breaks = breaks, labels = labels, include.lowest = TRUE)

# Convert the classified raster back to a RasterLayer object
classified_raster <- setValues(tot_f_vul_raster, classified_raster)

# Create a subtitle
subtitle <- "FVI = weigthed SOC + PHY Vulnerability Indicators"

intersect_grid <- st_intersection(grid, mahama_boundary)
```
# Final HFVI Visualization 

```{r HFVI Map}
# Set tmap mode to plotting
tmap_mode("plot")

# Define subtitle
subtitle <- "FVI = weighted SOC + PHY Vulnerability Indicators"

# Define color palette
palette <- "YlOrRd"

# Generate Jenks breaks
jenks_breaks <- classInt::classIntervals(values(tot_f_vul_raster), n = 4, style = "jenks", breaks = seq(0, 1))$brks

# Define labels
labels = c("None", "Low", "Medium", "High")

# Basemap
bbox_mahama
osm_map <- read_osm(bbox_mahama)

tm_f_vul_jenks <- tm_shape(osm_map) + 
  tm_rgb(alpha = 0.6) +  # Adjust transparency of the OSM map
  tm_layout(legend.position = c("left", "bottom"), 
            frame = FALSE) +  # Adjust legend position and remove frame
  tm_shape(tot_f_vul_raster) +
  tm_raster(palette = palette, 
            style = "fixed", 
            breaks = jenks_breaks, 
            labels = labels, 
            title = "Vulnerability Class", 
            alpha = 0.6, 
            legend.hist = TRUE) + 
  tm_shape(mahama_boundary) +
  tm_borders() +
  tm_scale_bar(position = c("left", "bottom")) +  # Add scale bar
  tm_compass(type="arrow", position=c(0, .2)) +  # Add compass

  tm_layout(title = "HFVI for the Mahama Refugee Camp",  # Add title
            title.position = c("left", "top"),  # Adjust title position
            title.size = 1.5,
            title.fontface = "bold",
            legend.hist.size = 1,
            legend.hist.width = 0.6) +  # Adjust title size
  
  tm_layout(legend.show = TRUE,
            legend.title.color = "black",
            legend.title.size = 0.9,
            legend.position = c("left", "bottom"),
            legend.text.color = "black")   # Adjust legend colors 


print(tm_f_vul_jenks)

tmap_save(tm_f_vul_jenks, "figs/HFVI_map.png", width = 7, height = 5)
```

```{r Mapping smoothed HFVI}

# Set tmap mode to plotting
tmap_mode("plot")

# Basemap
bbox_mahama
osm_map <- read_osm(bbox_mahama)

# Define breaks and labels for the legend
breaks <- seq(0, 1, by = 0.2)

tm_f_vul_smoothed <- tm_shape(osm_map) + 
  tm_rgb(alpha = 0.6) +  # Adjust transparency of the OSM map
  tm_layout(legend.position = c("left", "bottom"), frame = FALSE) +  # Adjust legend position and remove frame
  tm_shape(tot_f_vul_raster_smoothed) +
  tm_raster(palette = palette, 
            style = "cont", 
            breaks = breaks,
            title = "HFVI value", 
            alpha = 0.6, 
            legend.hist = TRUE) + 
  tm_shape(mahama_boundary) +
  tm_borders() +
  tm_scale_bar(position = c("left", "bottom")) +  # Add scale bar
  tm_compass(type="arrow", position=c(0, .2)) +  # Add compass
  
  tm_layout(title = "HFVI for the Mahama Refugee Camp",  # Add title
            title.position = c("left", "top"),  # Adjust title position
            title.size = 1.5,
            title.fontface = "bold") + # Adjust title size
  
  tm_layout(legend.show = TRUE,
            legend.title.color = "black",
            legend.title.size = 0.9,
            legend.position = c("left", "bottom"),
            legend.text.color = "black")   # Adjust legend colors 

tm_f_vul_smoothed

tmap_save(tm_f_vul_smoothed, "figs/HFVI_smoothed.png", width = 7, height = 5)

```

```{r Export raster}
writeRaster(tot_f_vul_raster, filename = "HFVI_f_raster.tif", overwrite=TRUE)
writeRaster(tot_f_vul_raster_smoothed, filename = "HFVI_f_raster_smoothed.tif", overwrite=TRUE)
writeRaster(f_overlayed_soc_smoothed, filename = "HFVI_soc.tif", overwrite=TRUE)
writeRaster(f_overlayed_phy_smoothed, filename = "HFVI_phy.tif", overwrite=TRUE)
```


# Spatial Analysis
Goals of spatial autocorrelation: Measuring the strength of spatial relationship and testing assumptions of independence and randomness

```{r Spatial Autocorrelation}
library(spdep)

# tot_vul is a SpatialPixelsDataFrame
tot_vul_spdf <- as(tot_f_vul, "SpatialPointsDataFrame")

# Extract raster values from tot_vul
tot_vul_values <- tot_vul_spdf@data[[1]]  

# Create neighbor list
nb <- cell2nb(nrow(tot_vul_spdf), ncol(tot_vul_spdf), type = "queen")

# Compute spatial weights matrix
w <- nb2listw(nb, style = "W")

# Calculate Moran's I
moran_I <- moran.test(tot_vul_values, w)

# Print Moran's I result
print(moran_I)

# Capture Moran's I result output
moran_I_output <- capture.output(print(moran_I))

# Write the output to a .txt file
writeLines(moran_I_output, "morans_I_output.txt")
```
Moran I statistic: This value measures the degree of spatial autocorrelation. It ranges from -1 to 1.
- A positive value (0.6431346025) moderate positive spatial autocorrelation, meaning similar values tend to cluster together.
- A negative value would indicate negative spatial autocorrelation, implying dissimilar values tend to cluster together.
- A value close to 0 suggests no spatial autocorrelation.

Expectation: This is the expected value of Moran's I under the assumption of spatial randomness. In this case, it's very close to zero, indicating no spatial autocorrelation under the null hypothesis of spatial randomness.

Variance: This measures the variance of the Moran I statistic under spatial randomness.

Standard Deviate and p-value: The standard deviate is a measure of how many standard deviations the observed Moran's I statistic is from the expected Moran's I under spatial randomness. The p-value indicates the significance of the Moran's I statistic.
Here, the Moran I statistic standard deviate is 27.937, and the p-value is very small (p-value < 2.2e-16), indicating strong evidence against the null hypothesis of spatial randomness.

Overall, the high Moran's I value and low p-value suggest significant positive spatial autocorrelation in the raster data, meaning similar values are clustered together in space.
--> see ETH slides: https://ethz.ch/content/dam/ethz/special-interest/baug/irl/plus-dam/documents/lehrveranstaltungen/msc/MCDA/Lecture4_criteria_maps.pdf
or https://geodacenter.github.io/workbook/5a_global_auto/lab5a.html#morans-i

```{r Moran's I scatter plot}
# Plot Moran's I scatter plot

mp <-moran.plot(tot_vul_values, w, zero.policy=attr(w, "zero.policy"), labels=NULL, xlab=NULL, ylab=NULL, quiet=NULL, plot=T)

if (require(ggplot2, quietly=TRUE)) {
  xname <- attr(mp, "xname")
  ggplot(mp, aes(x=x, y=wx)) + geom_point(shape=1) + 
    geom_smooth(formula=y ~ x, method="lm") + 
    geom_hline(yintercept=mean(mp$wx), lty=2) + 
    geom_vline(xintercept=mean(mp$x), lty=2) + theme_minimal() + 
    geom_point(data=mp[mp$is_inf,], aes(x=x, y=wx), shape=9) +
    geom_text(data=mp[mp$is_inf,], aes(x=x, y=wx, label=labels, vjust=1.5)) +
    xlab(xname) + ylab(paste0("Spatially lagged ", xname))
}
```

## Correlation Matrix

https://rdrr.io/cran/raster/man/layerStats.html

```{r Spatial correlation between vulnerability indicator rasters}

# stack the rasters 
stack_corr <- stack(soc1_raster, soc2_raster, soc3_raster, soc4_raster, phy1_raster, phy2_raster, phy3_raster, phy4_raster) 
# Rename the labels of the layers in the raster stack
names(stack_corr) <- c("SOC1", "SOC2", "SOC3", "SOC4", "PHY1", "PHY2", "PHY3", "PHY4")

stats_corr <- layerStats(stack_corr, 'pearson', na.rm=T)
corr_matrix <- stats_corr$'pearson correlation coefficient'
```

```{r Correlation matrix}
# Calculate significance of correlation: p-value 

# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

# matrix of the p-value of the correlation
p.mat <- cor.mtest(stack_corr)
head(p.mat[, 1:5])


col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

#png("figs/correlation_plot.png", width = 800, height = 800, res = 150)

corrplot(corr_matrix, method="color", col=col(200),  
         type="upper", 
         addCoef.col = "black", # Add coefficient of correlation
         font.cor = 2,
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         p.mat = p.mat, sig.level = 0.01, 
         # correlation coefficient on the principal diagonal
         diag=T, 
         # Adjust text properties for coefficients
         number.cex = 0.75, # Adjust the size of the coefficients
         mar = c(0,0,2,0) # Adjust margins
         )

#dev.off()
```
This can be used to explain the spatial pattern of the resulting values in the final HFVI map 

## Sensitivity Analysis 

https://zenodo.org/records/11044224 SpatMCDA package for spatial MDCA 
file:///Users/annikakunz/Documents/Uni/Master/Masterarbeit/R/SpatMCDA/vignettes/SpatMCDA.html 


```{r Mapping}
# SOC
# Stacking the rasters
rasters_soc <- raster::stack(
  soc1_raster,
  soc2_raster,
  soc3_raster,
  soc4_raster)

rasters_soc <- mask(rasters_soc, mahama_boundary)

weights_soc <- c(f_weight_soc1, f_weight_soc2, f_weight_soc3, f_weight_soc4) # Fuzzy AHP weights

soc_map <- wlc(rasters = rasters_soc,
               weight = weights_soc)

terra::writeRaster(soc_map, "./OAT/soc_map.tif", filetype = "GTiff", overwrite = TRUE)

# PHY
# Stacking the rasters
rasters_phy <- raster::stack(
  phy1_raster,
  phy2_raster,
  phy3_raster,
  phy4_raster)

rasters_phy <- mask(rasters_phy, mahama_boundary)

weights_phy <- c(f_weight_phy1, f_weight_phy2, f_weight_phy3, f_weight_phy4) # Fuzzy AHP weights

phy_map <- wlc(rasters = rasters_phy,
               weight = weights_phy)

terra::writeRaster(phy_map, "./OAT/phy_map.tif", filetype = "GTiff", overwrite = TRUE)
```

# OAT: One-at-a-time method 

The "oat" function generates two output result folders in the "output_dir": one folder named "dominant_factor" for outputting all rasters with adjusted weights, and another folder named "dominant_factor_difference" for summarizing the differences between the initial decision map and the rasters with adjusted weights. It also automatically plots all difference maps in the "Plots" window.

Ranges are adjusted to fit the FAHP range between wMIn and wMax:

SOC Fuzzy weights: 

indicator	wMin	wModal	wMax	Defuzzified_w	Norm_Defuzzified_w
SOC_1	0.2740	0.4033	0.6029	0.4267	0.3971
SOC_2	0.2234	0.3391	0.5249	0.3625	0.3373
SOC_3	0.1170	0.1836	0.3011	0.2005	0.1866
SOC_4	0.0443	0.0737	0.1368	0.0849	0.0790

```{r Fuzzy Weights Range: SOC}
# Adjusted Fuzzy Weights Range: SOC
soc1_min <- 0.2740
soc1_max <- 0.6029
soc2_min <- 0.2234
soc2_max <- 0.5249
soc3_min <- 0.1170
soc3_max <- 0.3011
soc4_min <- 0.0443
soc4_max <- 0.1368


# Calculate associated ranges of change: 
# SOC1
diff_to_min_soc1 <- (soc1_min - f_weight_soc1) * 100
diff_to_max_soc1 <- (soc1_max - f_weight_soc1) * 100

range_soc1 <- c(diff_to_min_soc1, diff_to_max_soc1)
range_soc1

# SOC2
diff_to_min_soc2 <- (soc2_min - f_weight_soc2) * 100
diff_to_max_soc2 <- (soc2_max - f_weight_soc2) * 100

range_soc2<- c(diff_to_min_soc2, diff_to_max_soc2)
range_soc2

# SOC3
diff_to_min_soc3 <- (soc3_min - f_weight_soc3) * 100
diff_to_max_soc3 <- (soc3_max - f_weight_soc3) * 100

range_soc3 <- c(diff_to_min_soc3, diff_to_max_soc3)
range_soc3

# SOC4
diff_to_min_soc4 <- (soc4_min - f_weight_soc4) * 100
diff_to_max_soc4 <- (soc4_max - f_weight_soc4) * 100

range_soc4 <- c(diff_to_min_soc4, diff_to_max_soc4)
range_soc4
```

PHY Fuzzy weights:

indicator	wMin	wModal	wMax	Defuzzified_w	Norm_Defuzzified_w
PHY_1	0.1513	0.2947	0.5697	0.3386	0.2962
PHY_2	0.1497	0.2582	0.4784	0.2954	0.2584
PHY_3	0.1670	0.3067	0.5762	0.3500	0.3062
PHY_4	0.0819	0.1406	0.2550	0.1592	0.1393

```{r Fuzzy Weights Range: PHY}
phy1_min <- 0.1513
phy1_max <- 0.5697
phy2_min <- 0.1497
phy2_max <- 0.4784
phy3_min <- 0.1670
phy3_max <- 0.5762
phy4_min <- 0.0819
phy4_max <- 0.2550

# Calculate associated ranges of change: 
# phy1
diff_to_min_phy1 <- (phy1_min - f_weight_phy1) * 100
diff_to_max_phy1 <- (phy1_max - f_weight_phy1) * 100

range_phy1 <- c(diff_to_min_phy1, diff_to_max_phy1)
range_phy1

# phy2
diff_to_min_phy2 <- (phy2_min - f_weight_phy2) * 100
diff_to_max_phy2 <- (phy2_max - f_weight_phy2) * 100

range_phy2<- c(diff_to_min_phy2, diff_to_max_phy2)
range_phy2

# phy3
diff_to_min_phy3 <- (phy3_min - f_weight_phy3) * 100
diff_to_max_phy3 <- (phy3_max - f_weight_phy3) * 100

range_phy3<- c(diff_to_min_phy3, diff_to_max_phy3)
range_phy3

# phy4
diff_to_min_phy4 <- (phy4_min - f_weight_phy4) * 100
diff_to_max_phy4 <- (phy4_max - f_weight_phy4) * 100

range_phy4<- c(diff_to_min_phy4, diff_to_max_phy4)
range_phy4
```

Next, we perform the sensitivity analysis of the risk map using One-At-a-Time.

The "oat" function generates two output result folders in the "output_dir": one folder named "dominant_factor" for outputting all rasters with adjusted weights, and another folder named "dominant_factor_difference" for summarizing the differences between the initial decision map and the rasters with adjusted weights. It also automatically plots all difference maps in the "Plots" window.

```{r OAT SOC1 / PHY1}

# SOC1
# Stacking the rasters
rasters_soc1 <- raster::stack(
  soc1_raster,
  soc2_raster,
  soc3_raster,
  soc4_raster)

# The order of the weights is aligned with the order of the rasters
weights <- c("soc1_raster" = f_weight_soc1, 
             "soc2_raster" = f_weight_soc2,
             "soc3_raster" = f_weight_soc3,
             "soc4_raster" = f_weight_soc4)

#Range of weight adjustment (%)
range <- range_soc1 # Ranges are adjusted to fit the FAHP range between wMIn and wMax 

#Step size of weight adjustment (%)
step <- 1

#All results are output in out_dir
oat(rasters = rasters_soc1,
    weights = weights,
    dominant_factor = "soc1_raster",
    range = range,
    step = step,
    plot = TRUE,
    output_dir = "./OAT")

# -------------------------------------------
# PHY1
# Stacking the rasters
rasters_phy1 <- raster::stack(
  phy1_raster,
  phy2_raster,
  phy3_raster,
  phy4_raster)

# The order of the weights is aligned with the order of the rasters
weights <- c("phy1_raster" = f_weight_phy1, 
             "phy2_raster" = f_weight_phy2,
             "phy3_raster" = f_weight_phy3,
             "phy4_raster" = f_weight_phy4)

#Range of weight adjustment (%)
range <- range_phy1

#Step size of weight adjustment (%)
step <- 1

#All results are output in out_dir
oat(rasters = rasters_phy1,
    weights = weights,
    dominant_factor = "phy1_raster",
    range = range,
    step = step,
    plot = TRUE,
    output_dir = "./OAT")

#Example: result_-20 =  Results at -20% change in weights
```

```{r OAT SOC2 / PHY2}
# Now change dominant factor -> amplifying SOC2 / PHY2 as main

# SOC2
# Stacking the rasters
rasters_soc2 <- raster::stack(
  soc2_raster,
  soc1_raster,
  soc3_raster,
  soc4_raster)

# The order of the weights is aligned with the order of the rasters
weights <- c("soc2_raster" = f_weight_soc2,
             "soc1_raster" = f_weight_soc1, 
             "soc3_raster" = f_weight_soc3,
             "soc4_raster" = f_weight_soc4)

#Range of weight adjustment (%)
range <- range_soc2

#Step size of weight adjustment (%)
step <- 1

#All results are output in out_dir
oat(rasters = rasters_soc2,
    weights = weights,
    dominant_factor = "soc2_raster",
    range = range,
    step = step,
    plot = TRUE,
    output_dir = "./OAT")

# ---------------------------------
# PHY2

# Stacking the rasters
rasters_phy2 <- raster::stack(
  phy2_raster,
  phy1_raster,
  phy3_raster,
  phy4_raster)

# The order of the weights is aligned with the order of the rasters
weights <- c("phy2_raster" = f_weight_phy2,
             "phy1_raster" = f_weight_phy1, 
             "phy3_raster" = f_weight_phy3,
             "phy4_raster" = f_weight_phy4)

#Range of weight adjustment (%)
range <- range_phy2

#Step size of weight adjustment (%)
step <- 1

#All results are output in out_dir
oat(rasters = rasters_phy2,
    weights = weights,
    dominant_factor = "phy2_raster",
    range = range,
    step = step,
    plot = TRUE,
    output_dir = "./OAT")
```

```{r OAT SOC3 / PHY3}
# Now change dominant factor -> amplifying SOC3 / PHY3 as main

# SOC
# Stacking the rasters
rasters_soc3 <- raster::stack(
  soc3_raster,
  soc2_raster,
  soc1_raster,
  soc4_raster)

# The order of the weights is aligned with the order of the rasters
weights <- c("soc3_raster" = f_weight_soc3,
             "soc2_raster" = f_weight_soc2,
             "soc1_raster" = f_weight_soc1, 
             "soc4_raster" = f_weight_soc4)

#Range of weight adjustment (%)
range <- range_soc3

#Step size of weight adjustment (%)
step <- 1

#All results are output in out_dir
oat(rasters = rasters_soc3,
    weights = weights,
    dominant_factor = "soc3_raster",
    range = range,
    step = step, 
    plot = TRUE,
    output_dir = "./OAT")

# ---------------------------------
# PHY3
# Stacking the rasters
rasters_phy3 <- raster::stack(
  phy3_raster,
  phy2_raster,
  phy1_raster,
  phy4_raster)

# The order of the weights is aligned with the order of the rasters
weights <- c("phy3_raster" = f_weight_phy3,
             "phy2_raster" = f_weight_phy2,
             "phy1_raster" = f_weight_phy1, 
             "phy4_raster" = f_weight_phy4)

#Range of weight adjustment (%)
range <- range_phy3

#Step size of weight adjustment (%)
step <- 1

#All results are output in out_dir
oat(rasters = rasters_phy3,
    weights = weights,
    dominant_factor = "phy3_raster",
    range = range,
    step = step, 
    plot = TRUE,
    output_dir = "./OAT")
```

```{r OAT SOC4 / PHY4}
# Now change dominant factor -> amplifying SOC4 /PHY4 as main

# SOC4
# Stacking the rasters
rasters_soc4 <- raster::stack(
  soc4_raster,
  soc3_raster,
  soc2_raster,
  soc1_raster)

# The order of the weights is aligned with the order of the rasters
weights <- c("soc4_raster" = f_weight_soc4,
             "soc3_raster" = f_weight_soc3,
             "soc2_raster" = f_weight_soc2,
             "soc1_raster" = f_weight_soc1)

#Range of weight adjustment (%)
range <- range_soc4

#Step size of weight adjustment (%)
step <- 1

#All results are output in out_dir
oat(rasters = rasters_soc4,
    weights = weights,
    dominant_factor = "soc4_raster",
    range = range,
    step = step,
    plot = TRUE,
    output_dir = "./OAT")

# -------------------------------
# SOC4

# Stacking the rasters
rasters_phy4 <- raster::stack(
  phy4_raster,
  phy3_raster,
  phy2_raster,
  phy1_raster)

# The order of the weights is aligned with the order of the rasters
weights <- c("phy4_raster" = f_weight_phy4,
             "phy3_raster" = f_weight_phy3,
             "phy2_raster" = f_weight_phy2,
             "phy1_raster" = f_weight_phy1)

#Range of weight adjustment (%)
range <- range_phy4

#Step size of weight adjustment (%)
step <- 1

#All results are output in out_dir
oat(rasters = rasters_phy4,
    weights = weights,
    dominant_factor = "phy4_raster",
    range = range,
    step = step,
    plot = TRUE,
    output_dir = "./OAT")
```

# Mean of absolute change rates (MACR)
Next, let's calculate the absolute average rate of change for all factors together!

```{r MACR SOC1 / PHY 1}
# SOC1

#Read the results of the One-Factor-a-Time of factor soc1
oat_rasters_soc1 <- raster::stack(dir("./OAT/soc1_raster",full.names = TRUE))

#Read the soc vulnerability map obtained with the weighted linear combination
wlc_raster_soc <- raster::raster("./OAT/soc_map.tif")

# Setting the output position
output_dir <- "./OAT/MACR"

#All results are output in out_dir
macr_soc1 <- macr(oat_rasters = oat_rasters_soc1, wlc_raster = wlc_raster_soc,
                    output_dir = output_dir, factor_name = "soc1_raster")

# -------------------------
# PHY1

#Read the results of the One-Factor-a-Time of factor phy1
oat_rasters_phy1 <- raster::stack(dir("./OAT/phy1_raster",full.names = TRUE))

#Read the phy vulnerability map obtained with the weighted linear combination
wlc_raster_phy <- raster::raster("./OAT/phy_map.tif")

# Setting the output position
output_dir <- "./OAT/MACR"

#All results are output in out_dir
macr_phy1 <- macr(oat_rasters = oat_rasters_phy1, wlc_raster = wlc_raster_phy,
                    output_dir = output_dir, factor_name = "phy1_raster")
```

Subsequently, let's calculate the mean of absolute change rates for the missing indicators. As shown in the process above, MACRs were calculated for each of the remaining 3 factors as shown in the process above.

```{r MACR}
# SOC

# SOC2
#Read the results of the One-Factor-a-Time of the factors
oat_rasters_soc2 <- raster::stack(dir("./OAT/soc2_raster",full.names = TRUE))

#All results are output in out_dir
macr_soc2 <- macr(oat_rasters = oat_rasters_soc2, wlc_raster = wlc_raster_soc,
                    output_dir = output_dir, factor_name = "soc2_raster")

# SOC3
oat_rasters_soc3 <- raster::stack(dir("./OAT/soc3_raster",full.names = TRUE))
macr_soc3 <- macr(oat_rasters = oat_rasters_soc3, wlc_raster = wlc_raster_soc,
                    output_dir = output_dir, factor_name = "soc3_raster")

# SOC4
oat_rasters_soc4 <- raster::stack(dir("./OAT/soc4_raster",full.names = TRUE))
macr_soc4 <- macr(oat_rasters = oat_rasters_soc4, wlc_raster = wlc_raster_soc,
                    output_dir = output_dir, factor_name = "soc4_raster")

# --------------------

# PHY

# phy2
#Read the results of the One-Factor-a-Time of the factors
oat_rasters_phy2 <- raster::stack(dir("./OAT/phy2_raster",full.names = TRUE))

#All results are output in out_dir
macr_phy2 <- macr(oat_rasters = oat_rasters_phy2, wlc_raster = wlc_raster_phy,
                    output_dir = output_dir, factor_name = "phy2_raster")

# phy3
oat_rasters_phy3 <- raster::stack(dir("./OAT/phy3_raster",full.names = TRUE))
macr_phy3 <- macr(oat_rasters = oat_rasters_phy3, wlc_raster = wlc_raster_phy,
                    output_dir = output_dir, factor_name = "phy3_raster")

# phy4
oat_rasters_phy4 <- raster::stack(dir("./OAT/phy4_raster",full.names = TRUE))
macr_phy4 <- macr(oat_rasters = oat_rasters_phy4, wlc_raster = wlc_raster_phy,
                    output_dir = output_dir, factor_name = "phy4_raster")
```

Let's plot the values of the MACRs for all factors together next!

```{r Read all MACR csv}

# SOC
# List of file paths: SOC
file_paths_soc <- c("~/Documents/Uni/Master/Masterarbeit/R/OAT/MACR/soc1_raster/soc1_raster_macr.csv",
                "~/Documents/Uni/Master/Masterarbeit/R/OAT/MACR/soc2_raster/soc2_raster_macr.csv",
                "~/Documents/Uni/Master/Masterarbeit/R/OAT/MACR/soc3_raster/soc3_raster_macr.csv",
                "~/Documents/Uni/Master/Masterarbeit/R/OAT/MACR/soc4_raster/soc4_raster_macr.csv")

# Function to process each file
process_file <- function(file_path) {
  df <- read.csv(file_path)  # Read the CSV file
  # Convert values in the column "CRW" to numeric based on the pattern
  df$CRW <- as.numeric(ifelse(str_detect(df$CRW, "^result_\\."), 
                              -as.numeric(str_replace(df$CRW, "^result_\\.", "")), 
                              as.numeric(str_replace(df$CRW, "^result_", ""))))
  # Rename columns
  names(df)[which(names(df) == "Factor_Name")] <- "Variable"
  names(df)[which(names(df) == "MACRs")] <- "Value"
  return(df)
}

# Apply the function to each file
processed_dfs_soc <- lapply(file_paths_soc, process_file)

# Combine all data frames into one
MACR_soc <- do.call(rbind, processed_dfs_soc)

# ----------------------
# PHY
# List of file paths: PHY
file_paths_phy <- c("~/Documents/Uni/Master/Masterarbeit/R/OAT/MACR/phy1_raster/phy1_raster_macr.csv",
                "~/Documents/Uni/Master/Masterarbeit/R/OAT/MACR/phy2_raster/phy2_raster_macr.csv",
                "~/Documents/Uni/Master/Masterarbeit/R/OAT/MACR/phy3_raster/phy3_raster_macr.csv",
                "~/Documents/Uni/Master/Masterarbeit/R/OAT/MACR/phy4_raster/phy4_raster_macr.csv")

# Function to process each file
process_file <- function(file_path) {
  df <- read.csv(file_path)  # Read the CSV file
  # Convert values in the column "CRW" to numeric based on the pattern
  df$CRW <- as.numeric(ifelse(str_detect(df$CRW, "^result_\\."), 
                              -as.numeric(str_replace(df$CRW, "^result_\\.", "")), 
                              as.numeric(str_replace(df$CRW, "^result_", ""))))
  # Rename columns
  names(df)[which(names(df) == "Factor_Name")] <- "Variable"
  names(df)[which(names(df) == "MACRs")] <- "Value"
  return(df)
}

# Apply the function to each file
processed_dfs_phy <- lapply(file_paths_phy, process_file)

# Combine all data frames into one
MACR_phy <- do.call(rbind, processed_dfs_phy)

```

```{r}
plot_macr(MACR_soc)
plot_macr(MACR_phy)
```

```{r}

# Custom labels for SOC variables
soc_custom_labels <- c(
  "soc1_raster" = "SOC1",
  "soc2_raster" = "SOC2",
  "soc3_raster" = "SOC3",
  "soc4_raster" = "SOC4"
)

phy_custom_labels <- c(
  "phy1_raster" = "PHY1",
  "phy2_raster" = "PHY2",
  "phy3_raster" = "PHY3",
  "phy4_raster" = "PHY4"
)

# Plot the data
macr_soc <- ggplot(MACR_soc, aes(x = CRW, y = Value, color = Variable)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "SOC Indicators",
       x = "Change rate of weights (CWR) [%]",
       y = " Mean Absolute Change Rates (MACR) [%]",
       color = "Variable") +
  theme(axis.text.x = element_text(hjust = 1)) +
  scale_color_manual(values = soc_custom_labels) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c"), labels = soc_custom_labels)

macr_phy <- ggplot(MACR_phy, aes(x = CRW, y = Value, color = Variable)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "PHY Indicators",
       x = "Change rate of weights (CWR) [%]",
       y = "Mean Absolute Change Rates (MACR) [%]",
       color = "Variable") +
  theme(axis.text.x = element_text(hjust = 1)) +
  scale_color_manual(values = c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c"), labels = phy_custom_labels)

# Arrange the plots next to each other
macr <- grid.arrange(macr_soc, macr_phy, ncol = 2)
ggsave("figs/macr_plots.png", plot = macr, width = 10, height = 5, dpi = 300)
```

Interpretation: as done in Chen et al., (2010)

# Uncertainty Map
Finally, let's put together all the  maps (all maps in all folders named "domain_factor" in the "oat" folder.) with adjusted weights. Let's calculate the uncertainty map.

First, copy all result tiffs from the individual indicators into the folder uncertainty_map_

```{r Uncertainty Map}

# SOC
# Create uncertainty map 
soc_results <- terra::rast(dir("./OAT/uncertainty_map_soc",full.names = TRUE))
uncertainty_soc <- uncertain(soc_results)

# Normalize
uncertainty_soc <- normalize(uncertainty_soc)

# Create the tm_plot object for the uncertainty map
soc_uncertainty <- tm_shape(uncertainty_soc) +
  tm_raster(palette = "plasma",
            style = "cont",
            title = "SOC Uncertainty")+
  tm_shape(mahama_boundary) +
  tm_borders() 

# Convert to RasterLayer
uncertainty_soc <- as(raster::raster(uncertainty_soc), "RasterLayer")

# Applying low-pass filter
un_soc_smoothed <- focal(uncertainty_soc, w = lowpass, fun=mean, na.rm = TRUE) %>%
  mask(., mahama_boundary)

# Normalize
un_soc_smoothed <- normalize(un_soc_smoothed)

# Plotting
# Create the tm_plot object for the uncertainty map
soc_uncertainty_smoothed <- tm_shape(un_soc_smoothed) +
  tm_raster(palette = "plasma",
            style = "cont",
            title = "SOC Uncertainty") +
  tm_shape(mahama_boundary) +
  tm_borders() 

# ---------------------
# PHY
# Create uncertainty map 
phy_results <- terra::rast(dir("./OAT/uncertainty_map_phy",full.names = TRUE))
uncertainty_phy <- uncertain(phy_results)

# Normalize
uncertainty_phy <- normalize(uncertainty_phy)

# Create the tm_plot object for the uncertainty map
phy_uncertainty <- tm_shape(uncertainty_phy) +
  tm_raster(palette = "plasma",
            style = "cont",
            title = "PHY Uncertainty") +
  tm_shape(mahama_boundary) +
  tm_borders() 

# Convert to RasterLayer
uncertainty_phy <- as(raster::raster(uncertainty_phy), "RasterLayer")

# Applying smoothing filter
un_phy_smoothed <- focal(uncertainty_phy, w = lowpass, fun=mean, na.rm = TRUE) %>%
  mask(., mahama_boundary)

# Normalize
un_phy_smoothed <- normalize(un_phy_smoothed)

# Plotting
# Create the tm_plot object for the uncertainty map
phy_uncertainty_smoothed <- tm_shape(un_phy_smoothed) +
  tm_raster(palette = "plasma",
            style = "cont",
            title = "PHY Uncertainty") +
  tm_shape(mahama_boundary) +
  tm_borders() 

# Plot the object
uncertainty_soc_phy <- tmap_arrange(soc_uncertainty, phy_uncertainty)
tmap_arrange(soc_uncertainty_smoothed, phy_uncertainty_smoothed)

tmap_save(uncertainty_soc_phy, "figs/uncertainty_soc_phy.png", width = 10, height = 5)
```

```{r Overlay Uncertainty Layers}
# Weighted linear overlay
overlay_unc <- (uncertainty_soc * 0.5) + (uncertainty_phy * 0.5)
overlay_unc <- normalize(overlay_unc)

smoothed_overlay_unc <- (un_soc_smoothed * 0.5) + (un_phy_smoothed * 0.5)
smoothed_overlay_unc <- normalize(smoothed_overlay_unc)
```

```{r Mapping combined uncertainty map}

tm_uncertainty <- tm_shape(overlay_unc) +
  tm_raster(palette = "plasma",
            style = "cont",
            title = "Index Uncertainty")

smoothed_overlay_unc <- mask(smoothed_overlay_unc, mahama_boundary)

tm_uncertainty_smoothed <- tm_shape(smoothed_overlay_unc) +
  tm_raster(palette = "plasma",
            style = "cont",
            title = "Index Uncertainty")

tmap_arrange(tm_uncertainty, tm_uncertainty_smoothed)
```

```{r Mapping Uncertainty}
# Set tmap mode to plotting
tmap_mode("plot")

# Basemap
bbox_mahama
osm_map <- read_osm(bbox_mahama)

# Define breaks and labels for the legend
breaks <- seq(0, 1, by = 0.2)

tm_unc <- tm_shape(osm_map) + 
  tm_rgb(alpha = 0.6) +  # Adjust transparency of the OSM map
  tm_layout(legend.position = c("left", "bottom"), frame = FALSE) +  # Adjust legend position and remove frame
  tm_shape(overlay_unc) +
  tm_raster(palette = "plasma", 
            style = "cont", 
            title = "Uncertainty Value", 
            alpha = 0.6, 
            legend.hist = TRUE) + 
  tm_shape(mahama_boundary) +
  tm_borders() +
  tm_scale_bar(position = c("left", "bottom")) +  # Add scale bar
  tm_compass(type="arrow", position=c(0, .2)) +  # Add compass
  
  tm_layout(title = "HFVI Weight Uncertainty Map",  # Add title
            title.position = c("left", "top"),  # Adjust title position
            title.size = 1.5,
            title.fontface = "bold",
            legend.hist.size = 1,
            legend.hist.width = 0.6) + # Adjust title size
  
  tm_layout(legend.show = TRUE,
            legend.title.color = "black",
            legend.title.size = 0.9,
            legend.position = c("left", "bottom"),
            legend.text.color = "black")   # Adjust legend colors 

tm_unc

# Smoothed 
# Basemap
bbox_mahama
osm_map <- read_osm(bbox_mahama)

# Define breaks and labels for the legend
breaks <- seq(0, 1, by = 0.2)

tm_unc_smoothed <- tm_shape(osm_map) + 
  tm_rgb(alpha = 0.6) +  # Adjust transparency of the OSM map
  tm_layout(legend.position = c("left", "bottom"), frame = FALSE) +  # Adjust legend position and remove frame
  tm_shape(smoothed_overlay_unc) +
  tm_raster(palette = "plasma", 
            style = "cont", 
            title = "Uncertainty Value", 
            alpha = 0.6, 
            legend.hist = TRUE) + 
  tm_shape(mahama_boundary) +
  tm_borders() +
  tm_scale_bar(position = c("left", "bottom")) +  # Add scale bar
  tm_compass(type="arrow", position=c(0, .2)) +  # Add compass
  
  tm_layout(title = "HFVI Weight Uncertainty Map - Smoothed",  # Add title
            title.position = c("left", "top"),  # Adjust title position
            title.size = 1.5,
            title.fontface = "bold",
            legend.hist.size = 1,
            legend.hist.width = 0.6) + # Adjust title size
  
  tm_layout(legend.show = TRUE,
            legend.title.color = "black",
            legend.title.size = 0.9,
            legend.position = c("left", "bottom"),
            legend.text.color = "black")   # Adjust legend colors 

tm_unc_smoothed
```


```{r Export and save raster}
writeRaster(overlay_unc, filename = "unc_raster.tif", overwrite=TRUE)
writeRaster(smoothed_overlay_unc, filename = "unc_raster_smoothed.tif", overwrite=TRUE)

tmap_save(tm_unc, "figs/unc_map.png", width = 7, height = 5)
tmap_save(tm_unc_smoothed, "figs/unc_map_smoothed.png", width = 7, height = 5)
```

```{r Final Maps}
tmap_arrange(tm_f_soc, soc_uncertainty)
tmap_arrange(tm_f_phy, phy_uncertainty)
tmap_arrange(tm_f_soc, soc_uncertainty, tm_f_phy, phy_uncertainty, nrow = 2)
tmap_arrange(tm_soc_smoothed, soc_uncertainty_smoothed, tm_phy_smoothed, phy_uncertainty_smoothed, nrow = 2)
tmap_arrange(tm_f_vul_smoothed, tm_uncertainty_smoothed, nrow = 1)
```

## Interpretation of WLC Vulnerability Mapping and OAT Weight Sensitivity Analysis 

*vulnerability Mapping with WLC:*
- The weighted linear combination (WLC) method is used to combine various raster layers representing vulnerability indicators using predefined weights.
- The resulting vulnerability represents the aggregated vulnerability level across all factors, with higher values indicating higher vulnerability.

*Sensitivity Analysis with OAT:*
- The One-At-a-Time (OAT) sensitivity analysis is performed by systematically adjusting the weights of individual factors while keeping the others constant. 
- Each factor is considered as the dominant factor, and its weight is adjusted within a specified range (depending on the fuzzy weights range) in steps of 1%.
- The output includes raster layers representing the vulnerability maps with adjusted weights for each factor (e.g., "result_-20" represents the vulnerability map with a 20% decrease in the weight of the dominant factor).
- The Mean of Absolute Change Rates (MACRs) quantifies the average rate of change across all factors, providing insights into the sensitivity of the overall vulnerability map to changes in individual factor weights.

*Uncertainty Map:*
The uncertainty map is generated by combining all vulnerability maps with adjusted weights from the OAT analysis. It represents the variability or uncertainty in the vulnerability assessment due to changes in factor weights.

*Interpretation:*
Interpretation of the results involves analyzing how changes in the weights of individual factors impact the overall vulnerability map. Factors with higher sensitivity (larger changes in vulnerability map output for small changes in weight) may be considered more influential in driving the overall vulnerability assessment. The uncertainty map provides insights into the stability and reliability of the vulnerability assessment, considering variations in factor weights.


*Further interpretation:*

Spatial Distribution: The uncertainty map will display spatial patterns indicating areas where changes in criteria weights have the most significant impact on the final result. Different regions on the map will show varying levels of uncertainty based on the sensitivity of the model to changes in criteria weights.

Magnitude of Uncertainty: The magnitude of uncertainty at each location on the map represents the degree of sensitivity of the model to changes in criteria weights. Higher uncertainty values indicate that small variations in criteria weights lead to significant changes in the final outcome, while lower uncertainty values suggest more robustness to changes in criteria weights.

Identifying Sensitive Areas: Areas with high uncertainty values indicate regions where decision-making might be more uncertain or where additional data or analysis may be necessary to make informed decisions. These areas could be considered 'hotspots' of uncertainty, requiring closer examination or further refinement of the MCA model.

Model Robustness: The overall pattern of the uncertainty map can provide insights into the robustness of your MCA model. If the uncertainty is uniformly distributed across the study area, it may suggest that the model is relatively insensitive to variations in criteria weights and thus more robust. Conversely, if certain regions exhibit significantly higher uncertainty, it may indicate areas where the model is less reliable and more sensitive to changes in criteria weights.

Decision Support: The uncertainty map can aid decision-makers by highlighting areas where uncertainty is high, allowing them to prioritize resources or further investigation in those regions. It also provides transparency regarding the reliability of the MCA model, helping stakeholders make more informed decisions based on the level of uncertainty associated with the results.

Overall, the final uncertainty map derived from the OAT sensitivity analysis provides valuable information about the robustness and reliability of your MCA model, allowing for more informed decision-making in spatial planning and resource management contexts.

```{r Map overlay of HFVI and uncertainties}

# Load your raster data
tot_f_vul_raster_smoothed
smoothed_overlay_unc

# Convert rasters to data frames
hfvi_df <- as.data.frame(tot_f_vul_raster_smoothed, xy = TRUE)
uncertainty_df <- as.data.frame(smoothed_overlay_unc, xy = TRUE)

# Rename columns for clarity
colnames(hfvi_df) <- c("x", "y", "HFVI")
colnames(uncertainty_df) <- c("x", "y", "Uncertainty")

# Convert data frames to sf objects
hfvi_sf <- st_as_sf(hfvi_df, coords = c("x", "y"), crs = st_crs(mahama_boundary))
uncertainty_sf <- st_as_sf(uncertainty_df, coords = c("x", "y"), crs = st_crs(mahama_boundary))

# Clip the sf objects to the mahama_boundary
hfvi_clipped <- st_intersection(hfvi_sf, mahama_boundary)
uncertainty_clipped <- st_intersection(uncertainty_sf, mahama_boundary)

# Convert back to data frames for ggplot
hfvi_clipped_df <- as.data.frame(st_coordinates(hfvi_clipped))
hfvi_clipped_df$HFVI <- hfvi_clipped$HFVI

uncertainty_clipped_df <- as.data.frame(st_coordinates(uncertainty_clipped))
uncertainty_clipped_df$Uncertainty <- uncertainty_clipped$Uncertainty

# Merge clipped data frames by coordinates
combined_df <- left_join(hfvi_clipped_df, uncertainty_clipped_df, by = c("X", "Y"))

# Normalize uncertainty values to range [0, 1] for saturation
combined_df <- combined_df %>%
  mutate(Uncertainty_normalized = scales::rescale(Uncertainty, to = c(0, 1)))

# Create the plot
ggplot(combined_df) +
  geom_raster(aes(x = X, y = Y, fill = HFVI, alpha = Uncertainty_normalized)) +
  scale_fill_gradientn(name = "HFVI", colors = brewer.pal(9, "YlOrRd")) +
  scale_alpha_continuous(name = "Uncertainty", range = c(0.1, 1)) + # Ensuring a suitable range for transparency
  coord_equal() +
  theme_void() +  # Remove the gray background and other default plot elements
  labs(title = "Overlay of HFVI and Uncertainty",
       subtitle = "Hue represents HFVI, Saturation represents Uncertainty")

```
