---
title: "FAHP analysis"
author: ""
date: "2024-04-02"
output: html_document
---

This code calculates the Fuzzy AHP weights based on the PCMs generated in the AHP.Rmd. It fuzzifies the inconsistency-corrected AHP PCMs and calculates individual and aggregates fuzzy weights and a final deffuzified value, which serves as final weight input for the HFVI calculation (HFVI.Rmd)


```{r setup, include=FALSE, warning = F}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# Install/load packages
## Default repository
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org" 
       options(repos = r)
})

check_pkg <- function(x)
  {
    if (!require(x, character.only = TRUE))
    {
      install.packages(x, dep = TRUE)
        if(!require(x, character.only = TRUE)) stop("Package not found")
    }
}

check_pkg("sf")
check_pkg("ggplot2")
check_pkg("plotly")
check_pkg("tidyverse")
check_pkg("leaflet")
check_pkg("dplyr")
check_pkg("XML")
check_pkg("mapview")
check_pkg("lubridate")
check_pkg("knitr")
check_pkg("xtable")
check_pkg("kableExtra")
check_pkg("gridExtra")
check_pkg("ahpsurvey")
check_pkg("FuzzyAHP")
check_pkg("readxl")
check_pkg("FuzzyAHP")
```

# Initialization of project paths
```{r init_paths, warning = FALSE, message = FALSE, fig.align = "center"}
library(here)
here::i_am("scripts/FAHP_analysis.Rmd")
knitr::opts_knit$set(root.dir = here::here())

dataFolder   <- here::here("data")   # Data folder
RFolder      <- here::here()         # RScript folder (i.e. where project file resides)
resultsFolder <- here::here("outputs/results")   # Figure folder
```


```{r Read data}
# AHP results: Data set containing all AHP questionnaire answers (rows) for each attribute
# Read CSV files from the 'data' folder
data_SOC <- read.csv(file.path(dataFolder, "ahp/ahp_soc.csv"), sep = ";", header = TRUE) %>%
  dplyr::select(-1)

data_PHY <- read.csv(file.path(dataFolder, "ahp/ahp_phy.csv"), sep = ";", header = TRUE) %>%
  dplyr::select(-1)

data_DOM <- read.csv(file.path(dataFolder, "ahp/ahp_dom.csv"), sep = ";", header = TRUE) %>%
  dplyr::select(-1)
```


```{r Assign attributes}
# Read SOC data
atts_soc <- c("SOC_1", "SOC_2", "SOC_3", "SOC_4") # Order needs to be correct
atts_phy <- c("PHY_1", "PHY_2", "PHY_3", "PHY_4")
atts_dom <- c("F_EXP", "S_SOC", "S_PHY")

atts_name_soc <- c("Population density", "Vulnerable Groups", "Facilities of social importance", "Land use")
atts_name_phy <- c("Shelter type", "Critical Infrastructure", "Facilities physical vulnerability", "Roads / Transport")
atts_name_dom <- c("Flood Exposure", "Social Susceptibility", "Physical Susceptibility")
```

```{r Use consistent filtered data}
# SOC: ID 6
data_SOC_filtered <- data_SOC[-c(2, 6), ]

# PHY: ID 6 & ID 11
data_PHY_filtered <- data_PHY[-c(6, 11), ]

# DOM: ID 6 & ID 11
data_DOM_filtered <- data_DOM[-c(6, 11), ]
```

# Fuzzy AHP 

Check articles Laarhoven and Pedrycz (1983) and Chang (1996)

Using the FuzzyAHP CRAN package

"Fuzzy logic is a very useful tool to transform qualitative assessments of a human to quantitative assessment by taking uncertainty into account. In a simple Analytical hierarchy process model, a decision maker has to do a weight scaling by carrying out comparisons in pair for his criteria and alternatives. He has to enter a definite number for his preferences, but in real world situation, a decision maker is not sure about the numerical values (weights) he should assign for his preferences. Fuzzy Analytical hierarchical process takes the real world uncertainty into account." https://hemi1984.medium.com/a-fuzzy-analytical-hierarchy-process-fahp-using-r-64c4c5a13b69 

The weights for criteria are, however, still determined from the pairwise comparison matrix. This approach towards AHP is common in Geosciences as well as other fields.

```{r Call documentation FuzzyAHP}
help(package="FuzzyAHP")
```

First, we have to construct a Fuzzy PCM based on the list of PCMs from the AHP ranking process. The FuzzyAHP packages includes a function which builds FPCMs based on AHP PCM inputs, resulting in Fuzzy Pairwise Comparison Matrix, calculated as minimum, geometric mean and maximum of each cell of all PCMs. 

```{r Load AHP PCMs}
pcm_soc <- data_SOC_filtered  %>% 
  ahp.mat(atts_soc, negconvert = T)

pcm_phy <- data_PHY_filtered %>% 
  ahp.mat(atts_phy, negconvert = T)

pcm_dom <- data_DOM_filtered %>%
  ahp.mat(atts_dom, negconvert = T)
```

The comparison matrices are fuzzified with respect to the fuzzy scale. Default fuzzy scale ranging from 1 to 9 with width of fuzzy number equal to 2 is provided.

```{r Fuzzification of PCMs: SOC}
# SOC

# Try with the first matrix in the list
pcm_soc_mat_1 <- as.matrix(pcm_soc[[1]])

comparisonMatrix <- pairwiseComparisonMatrix(pcm_soc_mat_1)

fuzzyComparisonMatrix <- fuzzyPairwiseComparisonMatrix(comparisonMatrix)
print(fuzzyComparisonMatrix)

# Create an empty list to store the fuzzy matrices
pcm_soc_fuzzy <- list()

# Loop through all elements from pcm_soc (1-9)
for (i in 1:9) {
  # Convert the current element to a matrix
  pcm_soc_mat <- as.matrix(pcm_soc[[i]])
  
  # Create a pairwise comparison matrix
  comparisonMatrix_soc <- pairwiseComparisonMatrix(pcm_soc_mat)
  
  # Fuzzify the pairwise comparison matrix
  fuzzyComparisonMatrix_soc <- fuzzyPairwiseComparisonMatrix(comparisonMatrix_soc)
  
  # Print the fuzzy comparison matrix
  print(fuzzyComparisonMatrix_soc)
  
  # Store the fuzzy comparison matrix in the list
  pcm_soc_fuzzy[[i]] <- fuzzyComparisonMatrix_soc
}
```

```{r Fuzzification of PCMs: PHY}
# PHY

# Create an empty list to store the fuzzy matrices
pcm_phy_fuzzy <- list()

# Loop through all elements from pcm_phy (1-9)
for (i in 1:9) {
  # Convert the current element to a matrix
  pcm_phy_mat <- as.matrix(pcm_phy[[i]])
  
  # Create a pairwise comparison matrix
  comparisonMatrix_phy <- pairwiseComparisonMatrix(pcm_phy_mat)
  
  # Fuzzify the pairwise comparison matrix
  fuzzyComparisonMatrix_phy <- fuzzyPairwiseComparisonMatrix(comparisonMatrix_phy)
  
  # Print the fuzzy comparison matrix
  print(fuzzyComparisonMatrix_phy)
  
  # Store the fuzzy comparison matrix in the list
  pcm_phy_fuzzy[[i]] <- fuzzyComparisonMatrix_phy
}
```

```{r Fuzzification of PCMs: DOM}
# DOM

# Create an empty list to store the fuzzy matrices
pcm_dom_fuzzy <- list()

# Loop through all elements from pcm_soc (1-9)
for (i in 1:9) {
  # Convert the current element to a matrix
  pcm_dom_mat <- as.matrix(pcm_dom[[i]])
  
  # Create a pairwise comparison matrix
  comparisonMatrix_dom <- pairwiseComparisonMatrix(pcm_dom_mat)
  
  # Fuzzify the pairwise comparison matrix
  fuzzyComparisonMatrix_dom <- fuzzyPairwiseComparisonMatrix(comparisonMatrix_dom)
  
  # Print the fuzzy comparison matrix
  print(fuzzyComparisonMatrix_dom)
  
  # Store the fuzzy comparison matrix in the list
  pcm_dom_fuzzy[[i]] <- fuzzyComparisonMatrix_dom
}

pcm_dom_fuzzy
```

```{r List PCMs for later fuzzification}
# SOC
# Create an empty list to store the pairwise comparison matrices
pcm_soc_comparison_matrices <- list()

# Loop through all elements from pcm_soc (1-9)
for (i in 1:length(pcm_soc)) {
  # Convert the current element to a matrix
  pcm_soc_mat <- as.matrix(pcm_soc[[i]])
  
  # Create a pairwise comparison matrix
  comparisonMatrix <- pairwiseComparisonMatrix(pcm_soc_mat)
  
  # Print the pairwise comparison matrix
  # print(comparisonMatrix)
  
  # Store the pairwise comparison matrix in the list
  pcm_soc_comparison_matrices[[i]] <- comparisonMatrix
}

# PHY
# Create an empty list to store the pairwise comparison matrices
pcm_phy_comparison_matrices <- list()

# Loop through all elements from pcm_phy (1-9)
for (i in 1:length(pcm_phy)) {
  # Convert the current element to a matrix
  pcm_phy_mat <- as.matrix(pcm_phy[[i]])
  
  # Create a pairwise comparison matrix
  comparisonMatrix <- pairwiseComparisonMatrix(pcm_phy_mat)
  
  # Print the pairwise comparison matrix
  # print(comparisonMatrix)
  
  # Store the pairwise comparison matrix in the list
  pcm_phy_comparison_matrices[[i]] <- comparisonMatrix
}
  
# DOM
# Create an empty list to store the pairwise comparison matrices
pcm_dom_comparison_matrices <- list()

# Loop through all elements from pcm_dom (1-9)
for (i in 1:length(pcm_dom)) {
  # Convert the current element to a matrix
  pcm_dom_mat <- as.matrix(pcm_dom[[i]])
  
  # Create a pairwise comparison matrix
  comparisonMatrix <- pairwiseComparisonMatrix(pcm_dom_mat)
  
  # Print the pairwise comparison matrix
  # print(comparisonMatrix)
  
  # Store the pairwise comparison matrix in the list
  pcm_dom_comparison_matrices[[i]] <- comparisonMatrix
}
```

This functions builds Fuzzy Pairwise Comparison Matrix based on list of Pairwise Comparison Matrices. The resulting Fuzzy Pairwise Comparison Matrix is calculated as minimum, geometric mean and maximum of each cell of all Pairwise Comparison Matrices in listOfMatrices.

```{r Fuzzy PCMs}
# SOC
# Turn PCMs into Fuzzy PCMs
fuzzy_soc <- buildFuzzyPairwiseComparisonMatrix(pcm_soc_comparison_matrices)

# PHY
# Turn PCMs into Fuzzy PCMs
fuzzy_phy <- buildFuzzyPairwiseComparisonMatrix(pcm_phy_comparison_matrices)

# DOM
# Turn PCMs into Fuzzy PCMs
fuzzy_dom <- buildFuzzyPairwiseComparisonMatrix(pcm_dom_comparison_matrices)
```

```{r Fuzzy PCM aggregated weights}
#SOC
# Calculate the weights of the fuzzy PCMs 
w_fuzzy_soc <- calculateWeights(fuzzy_soc)
# Calculate final fuzzy weight vector based on Chang (1996) to get the aggregated fuzzy weights including uncertainty in ranking process
calculate_weighting_vector(w_fuzzy_soc)

# PHY
# Calculate the weights of the fuzzy PCMs 
w_fuzzy_phy <- calculateWeights(fuzzy_phy)
# Calculate final fuzzy weight vector based on Chang (1996) to get the aggregated fuzzy weights including uncertainty in ranking process
calculate_weighting_vector(w_fuzzy_phy)

# DOM
# Calculate the weights of the fuzzy PCMs 
w_fuzzy_dom <- calculateWeights(fuzzy_dom)
# Calculate final fuzzy weight vector based on Chang (1996) to get the aggregated fuzzy weights including uncertainty in ranking process
agg_fuzz_w_dom <- calculate_weighting_vector(w_fuzzy_dom)
print("Aggregated Fuzzy weights - Vulnerability Domain:")
print(agg_fuzz_w_dom)
```
### ----

# Calculation of Fuzzy weights based on the procedure proposed by Al Fozaie (2022)

Apply Buckley’s geomean method to calculate the fuzzy weights. This involves multiplying the ﬁrst value of the triangular fuzzy scale of each column with one another to the root of 1/n, where n is the number of variables, and a is the ﬁrst integer corresponding to the triangular fuzzy value set a1, a2, a3 of proxy i; i.e., ri = [ai1 × ai2 × . . . × ain]1/n. Then the geomean is utilized to aggregate the values of these columns.


```{r Testing prodecure for a single FPCM}

# Start with testing this for the first FPCM of phy

# Define a function to calculate ri values for a given slot
calculate_ri <- function(slot_matrix, variable_names) {
  # Get the number of rows in the matrix
  n_rows <- nrow(slot_matrix)
  
  # Calculate the product of values for each row
  product_values <- apply(slot_matrix, 1, prod)
  
  # Calculate the root of 1/n
  root <- 1 / n_rows
  
  # Calculate ri for each row
  ri_values <- product_values ^ root
  
  # Create a data frame to store variable names and ri values
  ri_df <- data.frame(variable = variable_names, ri = ri_values)
  
  return(ri_df)
}

print(class(pcm_phy_fuzzy[[1]]))
print(pcm_phy_fuzzy[[1]])

# Get variable names
variable_names <- pcm_phy_fuzzy[[1]]@variableNames

# Calculate ri values for each slot
ri_fnMin <- calculate_ri(pcm_phy_fuzzy[[1]]@fnMin, variable_names)
ri_fnModal <- calculate_ri(pcm_phy_fuzzy[[1]]@fnModal, variable_names)
ri_fnMax <- calculate_ri(pcm_phy_fuzzy[[1]]@fnMax, variable_names)

# Merge ri values for fnMin, fnModal, and fnMax after the variable name
ri_merged <- merge(ri_fnMin, ri_fnModal, by = "variable")
ri_merged <- merge(ri_merged, ri_fnMax, by = "variable")

# Rename the ri values to rMin, rModal, and rMax
colnames(ri_merged) <- c("variable", "rMin", "rModal", "rMax")

# Display the merged ri values
print(ri_merged)

# Calculate the geometric mean of rMin, rModal, and rMax for each variable
ri_merged$geomean_ri <- apply(ri_merged[, -1], 1, function(x) exp(mean(log(x))))

# Display the updated dataframe
print(ri_merged)

kable(ri_merged) %>% kable_styling()
```
The next step is to calculate the fuzzy weights wi. In order to calculate the fuzzy weights, the researcher needs to calculate the sum of each column ri, and then multiply the inverse of ri to the value 

```{r Calculate fuzzy weight of test PCM}
# Calculate the sum of each ri column (rMin, rModal, rMax)
column_sums <- colSums(ri_merged[, c("rMin", "rModal", "rMax")])

# Display the column sums
print(column_sums)

# Calculate the inverse of the sum of each ri column (rMin, rModal, rMax)
inverse_sum <- 1 / column_sums

# Swap the positions of rMin and rMax in the inverse_sum vector
inverse_sum <- c(inverse_sum[3], inverse_sum[2], inverse_sum[1])

# Change variable names from "rMax" to "rMin" and vice versa
names(inverse_sum) <- c("rMin", "rModal", "rMax")

# Display the updated inverse_sum vector
print(inverse_sum)
```

```{r Fuzzy weights caculation for test PCM}
# Initialize an empty data frame to store the fuzzy weights
fuzzy_weights <- data.frame(variable = character(), wMin = numeric(), wModal = numeric(), wMax = numeric(), stringsAsFactors = FALSE)

# Define ri_cols variable to reference the columns in ri_merged corresponding to rMin, rModal, and rMax
ri_cols <- c("rMin", "rModal", "rMax")

# Calculate the fuzzy weights w for each variable
for (i in 1:nrow(ri_merged)) {
  row <- c(ri_merged[i, "variable"])
  for (j in 1:length(ri_cols)) {
    ri_value <- ri_merged[i, ri_cols[j]]
    inverse_sum_value <- inverse_sum[j]
    w_value <- round(inverse_sum_value * ri_value, 3)  # Round to 3 decimals
    row <- c(row, w_value)
  }
  # Append the row to the fuzzy weights data frame
  fuzzy_weights <- rbind(fuzzy_weights, row)
}

# Rename the columns
colnames(fuzzy_weights) <- c("variable", "wMin", "wModal", "wMax")

# Convert the fuzzy weights columns to numeric
fuzzy_weights[, 2:4] <- sapply(fuzzy_weights[, 2:4], as.numeric)

# Calculate the average of each row
fuzzy_weights$Defuzzified_w <- round(rowMeans(fuzzy_weights[, 2:4], na.rm = TRUE), 4)
# Calculate the average of each row
fuzzy_weights$Defuzzified_w_2 <- round((rowSums(fuzzy_weights[, 2:4], na.rm = TRUE) / 3), 4)

# ---

# Normalize defuzzified weights 

# Calculate the sum of defuzzified weights
sum_defuzzified <- sum(fuzzy_weights$Defuzzified_w)

# Normalize the defuzzified weights
fuzzy_weights$norm_defuzzified_w <- fuzzy_weights$Defuzzified_w / sum_defuzzified

# Display the updated fuzzy weights data frame
print(fuzzy_weights)

kable(fuzzy_weights) %>% kable_styling()
```

We are left with three columns that represent the fuzzy weights of the triangular fuzzy sets of the indicators.

The last step in the FAHP method is to defuzzy or normalize the weights. This includes calculating the average of the fuzzy weights for each proxy, i.e., defuzziﬁed. The
defuzziﬁed weights are the ﬁnal weights to be applied to the composite index after being subjected to the FAHP method.

# Looping this process to have all FPCMs of all experts

## SOC

```{r Fuzzy individual w: SOC}
# Initialize an empty list to store the results for each instance
results_list_soc <- list()

# Get variable names
variable_names <- pcm_soc_fuzzy[[1]]@variableNames

# Iterate over each instance in pcm_soc_fuzzy
for (instance_index in seq_along(pcm_soc_fuzzy)) {
  # Calculate ri values for each slot
  ri_fnMin <- calculate_ri(pcm_soc_fuzzy[[instance_index]]@fnMin, variable_names)
  ri_fnModal <- calculate_ri(pcm_soc_fuzzy[[instance_index]]@fnModal, variable_names)
  ri_fnMax <- calculate_ri(pcm_soc_fuzzy[[instance_index]]@fnMax, variable_names)
  
  # Merge ri values for fnMin, fnModal, and fnMax after the variable name
  ri_merged <- merge(ri_fnMin, ri_fnModal, by = "variable")
  ri_merged <- merge(ri_merged, ri_fnMax, by = "variable")
  
  # Rename the ri values to rMin, rModal, and rMax
  colnames(ri_merged) <- c("indicator", "rMin", "rModal", "rMax")
  
  # Calculate the geometric mean of rMin, rModal, and rMax for each variable
  ri_merged$geomean_ri <- apply(ri_merged[, -1], 1, function(x) exp(mean(log(x))))
  
  # Calculate the sum of each ri column (rMin, rModal, rMax)
  column_sums <- colSums(ri_merged[, c("rMin", "rModal", "rMax")])
  
  # Calculate the inverse of the sum of each ri column (rMin, rModal, rMax)
  inverse_sum <- 1 / column_sums
  
  # Swap the positions of rMin and rMax in the inverse_sum vector
  inverse_sum <- c(inverse_sum[3], inverse_sum[2], inverse_sum[1])
  
  # Change variable names from "rMax" to "rMin" and vice versa
  names(inverse_sum) <- c("rMin", "rModal", "rMax")
  
  # Initialize an empty data frame to store the fuzzy weights
  fuzzy_weights <- data.frame(indicator = character(), wMin = numeric(), wModal = numeric(), wMax = numeric(), stringsAsFactors = FALSE)
  
  # Calculate the fuzzy weights w for each indicator
  for (i in 1:nrow(ri_merged)) {
    row <- c(ri_merged[i, "indicator"])
    for (j in 1:length(ri_cols)) {
      ri_value <- ri_merged[i, ri_cols[j]]
      inverse_sum_value <- inverse_sum[j]
      w_value <- round(inverse_sum_value * ri_value, 3)  # Round to 3 decimals
      row <- c(row, w_value)
    }
    # Append the row to the fuzzy weights data frame
    fuzzy_weights <- rbind(fuzzy_weights, row)
  }
  
  # Rename the columns
  colnames(fuzzy_weights) <- c("indicator", "wMin", "wModal", "wMax")
  
  # Convert the fuzzy weights columns to numeric
  fuzzy_weights[, 2:4] <- sapply(fuzzy_weights[, 2:4], as.numeric)
  
  # Calculate the average of each row
  fuzzy_weights$Defuzzified_w <- round(rowMeans(fuzzy_weights[, 2:4], na.rm = TRUE), 4)
  
  # Store the results for this instance
  results_list_soc[[instance_index]] <- fuzzy_weights
}

# Display the results for each instance
for (i in seq_along(results_list_soc)) {
  cat("Results for instance", i, ":\n")
  print(results_list_soc[[i]])
  cat("\n")
}
```

Aggregated fuzzy weights: Now, calculating arithmetic mean over all the instances / FPCMs to get the the aggregated fuzzy weights

```{r Agggregated fuzzy weights SOC}
# Create an empty list to store the aggregated results
aggregated_results_soc <- list()

# Calculate the arithmetic mean for each column across all instances
for (col_name in c("wMin", "wModal", "wMax", "Defuzzified_w")) {
  # Extract the column values for all instances
  col_values <- sapply(results_list_soc, function(df) df[[col_name]])
  
  # Calculate the mean for each column
  mean_values <- round(rowMeans(col_values, na.rm = TRUE), 4)
  
  # Normalize the Defuzzified_w values to sum up to 1
  if (col_name == "Defuzzified_w") {
    normalized_values <- mean_values / sum(mean_values)
  } else {
    normalized_values <- mean_values
  }
  
  # Store the normalized values in the aggregated_results list
  aggregated_results_soc[[col_name]] <- mean_values
}

# Combine the mean values into a data frame
fuzzy_aggregated_soc <- data.frame(
  indicator = results_list_soc[[1]]$indicator, # Assuming all instances have the same indicators
  wMin = aggregated_results_soc[["wMin"]],
  wModal = aggregated_results_soc[["wModal"]],
  wMax = aggregated_results_soc[["wMax"]],
  Defuzzified_w = aggregated_results_soc[["Defuzzified_w"]],
  Norm_Defuzzified_w = round(normalized_values, 4)  # Add the normalized Defuzzified_w column
)

# Display the aggregated dataframe
print(fuzzy_aggregated_soc)

kable(fuzzy_aggregated_soc) %>% kable_styling()
```

## PHY

```{r Fuzzy individual w: PHY}
# Initialize an empty list to store the results for each instance
results_list_phy <- list()

# Get variable names
variable_names <- pcm_phy_fuzzy[[1]]@variableNames

# Iterate over each instance in pcm_phy_fuzzy
for (instance_index in seq_along(pcm_phy_fuzzy)) {
  # Calculate ri values for each slot
  ri_fnMin <- calculate_ri(pcm_phy_fuzzy[[instance_index]]@fnMin, variable_names)
  ri_fnModal <- calculate_ri(pcm_phy_fuzzy[[instance_index]]@fnModal, variable_names)
  ri_fnMax <- calculate_ri(pcm_phy_fuzzy[[instance_index]]@fnMax, variable_names)
  
  # Merge ri values for fnMin, fnModal, and fnMax after the variable name
  ri_merged <- merge(ri_fnMin, ri_fnModal, by = "variable")
  ri_merged <- merge(ri_merged, ri_fnMax, by = "variable")
  
  # Rename the ri values to rMin, rModal, and rMax
  colnames(ri_merged) <- c("indicator", "rMin", "rModal", "rMax")
  
  # Calculate the geometric mean of rMin, rModal, and rMax for each variable
  ri_merged$geomean_ri <- apply(ri_merged[, -1], 1, function(x) exp(mean(log(x))))
  
  # Calculate the sum of each ri column (rMin, rModal, rMax)
  column_sums <- colSums(ri_merged[, c("rMin", "rModal", "rMax")])
  
  # Calculate the inverse of the sum of each ri column (rMin, rModal, rMax)
  inverse_sum <- 1 / column_sums
  
  # Swap the positions of rMin and rMax in the inverse_sum vector
  inverse_sum <- c(inverse_sum[3], inverse_sum[2], inverse_sum[1])
  
  # Change variable names from "rMax" to "rMin" and vice versa
  names(inverse_sum) <- c("rMin", "rModal", "rMax")
  
  # Initialize an empty data frame to store the fuzzy weights
  fuzzy_weights <- data.frame(indicator = character(), wMin = numeric(), wModal = numeric(), wMax = numeric(), stringsAsFactors = FALSE)
  
  # Calculate the fuzzy weights w for each variable
  for (i in 1:nrow(ri_merged)) {
    row <- c(ri_merged[i, "indicator"])
    for (j in 1:length(ri_cols)) {
      ri_value <- ri_merged[i, ri_cols[j]]
      inverse_sum_value <- inverse_sum[j]
      w_value <- round(inverse_sum_value * ri_value, 3)  # Round to 3 decimals
      row <- c(row, w_value)
    }
    # Append the row to the fuzzy weights data frame
    fuzzy_weights <- rbind(fuzzy_weights, row)
  }
  
  # Rename the columns
  colnames(fuzzy_weights) <- c("indicator", "wMin", "wModal", "wMax")
  
  # Convert the fuzzy weights columns to numeric
  fuzzy_weights[, 2:4] <- sapply(fuzzy_weights[, 2:4], as.numeric)
  
  # Calculate the average of each row
  fuzzy_weights$Defuzzified_w <- round(rowMeans(fuzzy_weights[, 2:4], na.rm = TRUE), 4)
  
  # Store the results for this instance
  results_list_phy[[instance_index]] <- fuzzy_weights
}

# Display the results for each instance
for (i in seq_along(results_list_phy)) {
  cat("Results for instance", i, ":\n")
  print(results_list_phy[[i]])
  cat("\n")
}
```
```{r Agggregated fuzzy weights PHY}
# Create an empty list to store the aggregated results
aggregated_results_phy <- list()

# Calculate the arithmetic mean for each column across all instances
for (col_name in c("wMin", "wModal", "wMax", "Defuzzified_w")) {
  # Extract the column values for all instances
  col_values <- sapply(results_list_phy, function(df) df[[col_name]])
  
  # Calculate the mean for each column
  mean_values <- round(rowMeans(col_values, na.rm = TRUE), 4)
  
  # Normalize the Defuzzified_w values to sum up to 1
  if (col_name == "Defuzzified_w") {
    normalized_values <- mean_values / sum(mean_values)
  } else {
    normalized_values <- mean_values
  }
  
  # Store the normalized values in the aggregated_results list
  aggregated_results_phy[[col_name]] <- mean_values
}

# Combine the mean values into a data frame
fuzzy_aggregated_phy <- data.frame(
  indicator = results_list_phy[[1]]$indicator, # Assuming all instances have the same indicators
  wMin = aggregated_results_phy[["wMin"]],
  wModal = aggregated_results_phy[["wModal"]],
  wMax = aggregated_results_phy[["wMax"]],
  Defuzzified_w = aggregated_results_phy[["Defuzzified_w"]],
  Norm_Defuzzified_w = round(normalized_values, 4)  # Add the normalized Defuzzified_w column
)

# Display the aggregated dataframe
print(fuzzy_aggregated_phy)

kable(fuzzy_aggregated_phy) %>% kable_styling()
```

## DOM

```{r DOM test}

# Start with testing this for the first FPCM of dom

# Define a function to calculate ri values for a given slot
calculate_ri <- function(slot_matrix, variable_names) {
  # Get the number of rows in the matrix
  n_rows <- nrow(slot_matrix)
  
  # Calculate the product of values for each row
  product_values <- apply(slot_matrix, 1, prod)
  
  # Calculate the root of 1/n
  root <- 1 / n_rows
  
  # Calculate ri for each row
  ri_values <- product_values ^ root
  
  # Create a data frame to store variable names and ri values
  ri_df <- data.frame(variable = variable_names, ri = ri_values)
  
  return(ri_df)
}

# Get variable names
variable_names <- pcm_dom_fuzzy[[1]]@variableNames

# Calculate ri values for each slot
ri_fnMin <- calculate_ri(pcm_dom_fuzzy[[1]]@fnMin, variable_names)
ri_fnModal <- calculate_ri(pcm_dom_fuzzy[[1]]@fnModal, variable_names)
ri_fnMax <- calculate_ri(pcm_dom_fuzzy[[1]]@fnMax, variable_names)

# Merge ri values for fnMin, fnModal, and fnMax after the variable name
ri_merged <- merge(ri_fnMin, ri_fnModal, by = "variable")
ri_merged <- merge(ri_merged, ri_fnMax, by = "variable")

# Rename the ri values to rMin, rModal, and rMax
colnames(ri_merged) <- c("indicator", "rMin", "rModal", "rMax")

# Display the merged ri values
print(ri_merged)

# Calculate the geometric mean of rMin, rModal, and rMax for each variable
ri_merged$geomean_ri <- apply(ri_merged[, -1], 1, function(x) exp(mean(log(x))))

# Display the updated dataframe
print(ri_merged)

kable(ri_merged) %>% kable_styling()
```


```{r Fuzzy individual w: DOM}
# Initialize an empty list to store the results for each instance
results_list_dom <- list()

# Iterate over each instance in pcm_dom_fuzzy
for (instance_index in seq_along(pcm_dom_fuzzy)) {
  # Calculate ri values for each slot
  ri_fnMin <- calculate_ri(pcm_dom_fuzzy[[instance_index]]@fnMin, variable_names)
  ri_fnModal <- calculate_ri(pcm_dom_fuzzy[[instance_index]]@fnModal, variable_names)
  ri_fnMax <- calculate_ri(pcm_dom_fuzzy[[instance_index]]@fnMax, variable_names)
  
  # Merge ri values for fnMin, fnModal, and fnMax after the variable name
  ri_merged <- merge(ri_fnMin, ri_fnModal, by = "variable")
  ri_merged <- merge(ri_merged, ri_fnMax, by = "variable")
  
  # Rename the ri values to rMin, rModal, and rMax
  colnames(ri_merged) <- c("indicator", "rMin", "rModal", "rMax")
  
  # Normalize the modal values
  ri_merged$rModal <- ri_merged$rModal / sum(ri_merged$rModal)
  
  # Calculate the geometric mean of rMin, rModal, and rMax for each variable
  ri_merged$geomean_ri <- apply(ri_merged[, -1], 1, function(x) exp(mean(log(x))))
  
  # Calculate the sum of each ri column (rMin, rModal, rMax)
  column_sums <- colSums(ri_merged[, c("rMin", "rModal", "rMax")])
  
  # Calculate the inverse of the sum of each ri column (rMin, rModal, rMax)
  inverse_sum <- 1 / column_sums
  
  # Swap the positions of rMin and rMax in the inverse_sum vector
  inverse_sum <- c(inverse_sum[3], inverse_sum[2], inverse_sum[1])
  
  # Change variable names from "rMax" to "rMin" and vice versa
  names(inverse_sum) <- c("rMin", "rModal", "rMax")
  
  # Initialize an empty data frame to store the fuzzy weights
  fuzzy_weights <- data.frame(indicator = character(), wMin = numeric(), wModal = numeric(), wMax = numeric(), stringsAsFactors = FALSE)
  
  # Calculate the fuzzy weights w for each variable
  for (i in 1:nrow(ri_merged)) {
    row <- c(ri_merged[i, "indicator"])
    for (j in 1:length(ri_cols)) {
      ri_value <- ri_merged[i, ri_cols[j]]
      inverse_sum_value <- inverse_sum[j]
      w_value <- round(inverse_sum_value * ri_value, 3)  # Round to 3 decimals
      row <- c(row, w_value)
    }
    # Append the row to the fuzzy weights data frame
    fuzzy_weights <- rbind(fuzzy_weights, row)
  }
  
  # Rename the columns
  colnames(fuzzy_weights) <- c("indicator", "wMin", "wModal", "wMax")
  
  # Convert the fuzzy weights columns to numeric
  fuzzy_weights[, 2:4] <- sapply(fuzzy_weights[, 2:4], as.numeric)
  
  # Calculate the average of each row
  fuzzy_weights$Defuzzified_w <- round(rowMeans(fuzzy_weights[, 2:4], na.rm = TRUE), 4)
  
  # Store the results for this instance
  results_list_dom[[instance_index]] <- fuzzy_weights
}

# Display the results for each instance
for (i in seq_along(results_list_dom)) {
  cat("Results for instance", i, ":\n")
  print(results_list_dom[[i]])
  cat("\n")
}
```

```{r Agggregated fuzzy weights DOM}
# Create an empty list to store the aggregated results
aggregated_results_dom <- list()

# Calculate the arithmetic mean for each column across all instances
for (col_name in c("wMin", "wModal", "wMax", "Defuzzified_w")) {
  # Extract the column values for all instances
  col_values <- sapply(results_list_dom, function(df) df[[col_name]])
  
  # Calculate the mean for each column
  mean_values <- round(rowMeans(col_values, na.rm = TRUE), 4)
  
  # Normalize the Defuzzified_w values to sum up to 1
  if (col_name == "Defuzzified_w") {
    normalized_values <- mean_values / sum(mean_values)
  } else {
    normalized_values <- mean_values
  }
  
  # Store the normalized values in the aggregated_results list
  aggregated_results_dom[[col_name]] <- mean_values
}

# Combine the mean values into a data frame
fuzzy_aggregated_dom <- data.frame(
  indicator = results_list_dom[[1]]$indicator, # Assuming all instances have the same indicators
  wMin = aggregated_results_dom[["wMin"]],
  wModal = aggregated_results_dom[["wModal"]],
  wMax = aggregated_results_dom[["wMax"]],
  Defuzzified_w = aggregated_results_dom[["Defuzzified_w"]],
  Norm_Defuzzified_w = round(normalized_values, 4)  # Add the normalized Defuzzified_w column
)

# Display the aggregated dataframe
print(fuzzy_aggregated_dom)

kable(fuzzy_aggregated_dom) %>% kable_styling()
```


```{r Visualize fuzzy weights distribution}
# Plot the interval
f_soc_plt <- ggplot(fuzzy_aggregated_soc, aes(x = indicator)) +
  geom_errorbar(aes(ymin = wMin, ymax = wMax, color = "Interval"), width = 0.2, show.legend = TRUE, color = "#808080") +  # Plot intervals as error bars
  geom_point(aes(y = wModal, color = "Modal"), size = 3, show.legend = TRUE, shape = 10) +  # Plot modal values as points
  geom_point(aes(y = Defuzzified_w, color = "Defuzzified"), size = 3, show.legend = TRUE, shape = 10) +  # Plot defuzzified values as points
  labs(title = "", subtitle = "SOC Indicators",
       x = "",
       y = "Relative Weight",
       color = "Aggregated fuzzy weight:") +  
  scale_color_manual(values = c("#3366FF", "#33CC33"), labels = c("wModal", "wDefuzzified")) +  # Set color legend labels
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") +  # Rotate x-axis labels for better readability 
  scale_x_discrete(labels = c("SOC1", "SOC2", "SOC3", "SOC4")) +  # Change x-axis labels
  ylim(0.0, 0.61)  # Set y-axis limits

f_soc_plt

# Plot the interval
f_phy_plt <- ggplot(fuzzy_aggregated_phy, aes(x = indicator)) +
  geom_errorbar(aes(ymin = wMin, ymax = wMax, color = "Interval"), width = 0.2, show.legend = TRUE, color = "#808080") +  # Plot intervals as error bars
  geom_point(aes(y = wModal, color = "Modal"), size = 3, show.legend = TRUE, shape = 10) +  # Plot modal values as points
  geom_point(aes(y = Defuzzified_w, color = "Defuzzified"), size = 3, show.legend = TRUE, shape = 10) +  # Plot defuzzified values as points
  labs(title = "", subtitle = "PHY Indicators",
       x = "",
       y = "",
       color = "Aggregated fuzzy weight:") +  
  scale_color_manual(values = c("#3366FF", "#33CC33"), labels = c("wModal", "wDefuzzified")) +  # Set color legend labels
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") +  # Rotate x-axis labels for better readability
  scale_x_discrete(labels = c("PHY1", "PHY2", "PHY3", "PHY4")) +  # Change x-axis labels
  ylim(0.0, 0.6)  # Set y-axis limits

f_phy_plt

# Plot the interval
f_dom_plt <- ggplot(fuzzy_aggregated_dom, aes(x = indicator)) +
  geom_errorbar(aes(ymin = wMin, ymax = wMax, color = "Interval"), width = 0.2, show.legend = TRUE, color = "#808080") +  # Plot intervals as error bars
  geom_point(aes(y = wModal, color = "Modal"), size = 3, show.legend = TRUE, shape = 10) +  # Plot modal values as points
  geom_point(aes(y = Defuzzified_w, color = "Defuzzified"), size = 3, show.legend = TRUE, shape = 10) +  # Plot defuzzified values as points
  labs(title = "Fuzzy Weight Distribution for Vulnerability Indicators",
       x = "",
       y = "Relative Weight",
       color = "Aggregated fuzzy weight:") +  
  scale_color_manual(values = c("#3366FF", "#33CC33"), labels = c("wModal", "wDefuzzified")) +  # Set color legend labels
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top")  # Rotate x-axis labels for better readability

f_dom_plt
```
```{r save plots}
# Save plots to the "results" folder
ggsave(file.path(resultsFolder, "fahp/f_w_soc.png"), f_soc_plt , width = 12, height = 4)
ggsave(file.path(resultsFolder, "fahp/f_w_phy.png"), f_phy_plt, width = 12, height = 4)

# Arrange the two plots vertically
f_w <- grid.arrange(f_soc_plt, f_phy_plt, ncol = 2)

ggsave(file.path(resultsFolder, "fahp/f_w.png"), f_w, width = 10, height = 5)
```

